---
title: "R Notebook"
output: html_notebook
---


Set working directory
```{r}
setwd("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains")
```
Set Java directory
```{r}
Sys.setenv(JAVA_HOME='C:/Program Files/Java/jre1.8.0_361') # for 64-bit version
```



#Install libraries
```{r}
install.packages("multcompView")
install.packages("lmtest")
install.packages("relaimpo")
install.packages("ggplot2")
install.packages("ggbeeswarm")
install.packages("cati")
install.packages("ecodist")
install.packages("flextable")
install.packages("Matrix")
install.packages("lubridate")
install.packages("tidyverse")
install.packages("raster")
install.packages("ggpubr")
install.packages("rstatix")
install.packages("xlsx")
install.packages("BiodiversityR") #includes vegan
#install.packages("vegan")
install.packages("skimr")
install.packages("nlme")
install.packages("ape")
install.packages("Hmisc")
install.packages("devtools")
install.packages("vctrs")
devtools::install_github("dustinfife/flexplot")
devtools::install_github("hiweller/colordistance", build_vignettes=TRUE)
install.packages("MCMCglmm")
install.packages("sjPlot")
install.packages("glmmTMB")
install.packages("MuMIn")
install.packages("ade4")
install.packages("MDMR")
install.packages("gplots")
install.packages("RColorBrewer")
install.packages("svglite")
install.packages("ggfortify")
devtools::install_github("kylebittinger/usedist")
install.packages("see")
devtools::install_github("psyteachr/introdataviz")
install.packages("ggrain")
install.packages("ftExtra")
```


# Load libraries
```{r}
library(multcompView)
library(lmtest)
library(relaimpo)
#library(cati)
library(ecodist)
library(flextable)
#library(BiodiversityR)
library(lubridate)
library(tidyverse)
library(raster)
library(ggpubr)
library(rstatix)
library(xlsx)
library(skimr)
library(vegan)
library(stringr)
library(nlme)
library(ape)
library(Hmisc)
library(lme4)
library(flexplot)
library(MASS)
library(colordistance)
library(MCMCglmm)
library(sjPlot)
library(glmmTMB)
library(car)
library(boot)
library(MuMIn)
library(ade4)
library(cluster)
library(MDMR)
library(gplots)
library(RColorBrewer)
library(svglite)
library(ggfortify)
library(usedist)
library(ggbeeswarm)
library(see)
library(introdataviz)
library(ggrain)
library(ftExtra)
```


#Read data

The following section reads and correctly prepares all my data.

```{r}
prelim.data <- read.csv("C:/Users/HeraD/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/Experiment1_prelimdata_20240315.csv")

prelim.data <- prelim.data[-325,] #remove empty row
prelim.data <- prelim.data[-325,] #remove empty rows
```

Convert columns into factors
```{r}
# Species
prelim.data$Species.Code <- factor(prelim.data$Species.Code)
# Location code
prelim.data$Location_code <- factor(prelim.data$Location_code)
# Position
prelim.data$Experiment1Positition <- factor(prelim.data$Experiment1Positition)
# Sample source (replicate ID)
prelim.data$Sample.source <- factor(prelim.data$Sample.source)
# Wild / cultivated type
prelim.data$Cultivated.wild <- factor(prelim.data$Cultivated.wild)
# Clade
#prelim.data$Clade <- factor(prelim.data$Clade)
# Species
prelim.data$Species <- factor(prelim.data$Species)
# Morphotype
prelim.data$Morphotype <- factor(prelim.data$Morphotype)
# Inoculation date
prelim.data$Date <- as.Date(prelim.data$Date)
# All dates
prelim.data$SubstrateColonisationT1_date <- as.Date(prelim.data$SubstrateColonisationT1_date)
prelim.data$SubstrateColonisationT2_date <- as.Date(prelim.data$SubstrateColonisationT2_date)
prelim.data$Flush1PinningDate <- as.Date(prelim.data$Flush1PinningDate)
prelim.data$Flush1PinningDate_Away <- as.Date(prelim.data$Flush1PinningDate_Away)
prelim.data$Flush1HarvestDate <- as.Date(prelim.data$Flush1HarvestDate)
prelim.data$Flush2PinningDate <- as.Date(prelim.data$Flush2PinningDate)
prelim.data$Flush2HarvestDate <- as.Date(prelim.data$Flush2HarvestDate)
#prelim.data$Flush3PinningDate <- as.Date(prelim.data$Flush3PinningDate)
prelim.data$Flush3HarvestDate <- as.Date(prelim.data$Flush3HarvestDate)
# Other factors
prelim.data$Flush1Fruitbody_size <- factor(prelim.data$Flush1Fruitbody_size)
prelim.data$Flush2Fruitbody_size <- factor(prelim.data$Flush2Fruitbody_size)
prelim.data$Flush3Fruitbody_size <- factor(prelim.data$Flush3Fruitbody_size)
prelim.data$Flush1Fruitbody_captexture <- factor(prelim.data$Flush1Fruitbody_captexture)
prelim.data$Flush2Fruitbody_captexture <- factor(prelim.data$Flush2Fruitbody_captexture)
prelim.data$Flush3Fruitbody_captexture <- factor(prelim.data$Flush3Fruitbody_captexture)
prelim.data$Sporeprint_colour <- factor(prelim.data$Sporeprint_colour)
prelim.data$ImmatureHarvestWetWEight <- as.numeric(prelim.data$ImmatureHarvestWetWEight)
prelim.data$SubstrateEndWeightWet_raw <- as.numeric(prelim.data$SubstrateEndWeightWet_raw)

```

##Calculating time-based traits
```{r}
#Time from inoculation until fruiting
prelim.data$Flush1Time_to_Fruiting <- as.numeric(prelim.data$Flush1PinningDate - prelim.data$Date)

#Time from fruiting until harvest
prelim.data$Flush1Time_to_Mature <- as.numeric(prelim.data$Flush1HarvestDate - prelim.data$Flush1PinningDate)

```


##Calculating dry weight and water content
```{r}
# First flush fruiting body dry weight
prelim.data$Flush1Fruitbody_total_dryweight <- (prelim.data$Flush1Fruitbody_weight_dry/(prelim.data$Flush1Fruitbody_weight_wet_substrateremoved - prelim.data$Sporeprint_subsampl_weight) * prelim.data$Flush1Fruitbody_weight_wet_substrateremoved)

# First flush fruiting body water content
prelim.data$Flush1Fruitbody_water_content <- 1-(prelim.data$Flush1Fruitbody_weight_dry/(prelim.data$Flush1Fruitbody_weight_wet_substrateremoved - prelim.data$Sporeprint_subsampl_weight))

# Second flush fruiting body dry weight
prelim.data$Flush2Fruitbody_total_dryweight <- prelim.data$Flush2Fruitbody_weight_dry

# Second flush fruiting body water content
prelim.data$Flush2Fruitbody_water_content <- 1-(prelim.data$Flush2Fruitbody_weight_dry/(prelim.data$Flush2Fruitbody_weight_wet_substrateremoved))

# Third flush fruiting body dry weight
prelim.data$Flush3Fruitbody_total_dryweight <- prelim.data$Flush3Fruitbody_weight_dry

# Third flush fruiting body water content
prelim.data$Flush3Fruitbody_water_content <- 1-(prelim.data$Flush3Fruitbody_weight_dry/(prelim.data$Flush3Fruitbody_weight_wet_substrateremoved))

# Fourth flush fruiting body dry weight
prelim.data$Flush4Fruitbody_total_dryweight <- prelim.data$Flush4Fruitbody_weight_dry

# Fourth flush fruiting body water content
prelim.data$Flush4Fruitbody_water_content <- 1-(prelim.data$Flush4Fruitbody_weight_dry/(prelim.data$Flush4Fruitbody_weight_wet_substrateremoved))

# Total biomass produced
allweights <- c("Flush1Fruitbody_total_dryweight", "Flush2Fruitbody_total_dryweight", "Flush3Fruitbody_total_dryweight", "Flush4Fruitbody_total_dryweight", "ImmatureHarvestDryWeight", "AsexualGrowthWeightDry")

prelim.data$TotalBiomassProducedDryWeight <- apply(prelim.data[, allweights], 1, sum, na.rm = T) #for each row, sum up all weights of fruitbodies, removing NAs. because this returns 0s where there were NAs in all rows, the following function replaces all 0s with NAs
prelim.data$TotalBiomassProducedDryWeight <- replace(prelim.data$TotalBiomassProducedDryWeight, prelim.data$TotalBiomassProducedDryWeight == 0, NA)

# Immature fruiting body dry weight (end of experiment)
#prelim.data$ImmatureHarvestDryWeight 

# Aborted Australis fruiting body dry weight (end of experiment)
#prelim.data$AsexualGrowthWeightDry 

## Substrate mass and substrate mass loss
prelim.data[prelim.data$ImmatureHarvestWetWeight == "at MWLR"] <- NA # need to sort out this column with missing values
prelim.data$SubstrateMassInitial <- 231 # 200 g of sawdust and 31g of grainspawn
PaperBagWeightDry <- 20.87
PlasticBagWeightDry <- 11.23
prelim.data$SubstrateMassInitial[which(prelim.data$Sample.source == "Control")] <- 200 # correcting weight of Control

# replace NAs with zeroes to calculate actual mass loss, subtracting any other data, because R can't handle NAs for the next step of calculating the percentage of mass loss
prelim.data$SubstrateMassLoss <- 
  prelim.data$SubstrateMassInitial -
  (prelim.data$SubstrateEndWeightDry_raw - PaperBagWeightDry - PlasticBagWeightDry) -
  (
    (ifelse(is.na(prelim.data$Flush1Fruitbody_weight_wet), 0, prelim.data$Flush1Fruitbody_weight_wet) -
       ifelse(is.na(prelim.data$Flush1Fruitbody_weight_wet_substrateremoved), 0, prelim.data$Flush1Fruitbody_weight_wet_substrateremoved)) +
      (ifelse(is.na(prelim.data$Flush2Fruitbody_weight_wet), 0, prelim.data$Flush2Fruitbody_weight_wet) -
       ifelse(is.na(prelim.data$Flush2Fruitbody_weight_wet_substrateremoved), 0, prelim.data$Flush2Fruitbody_weight_wet_substrateremoved)) +
      (ifelse(is.na(prelim.data$Flush3Fruitbody_weight_wet), 0, prelim.data$Flush3Fruitbody_weight_wet) -
       ifelse(is.na(prelim.data$Flush3Fruitbody_weight_wet_substrateremoved), 0, prelim.data$Flush3Fruitbody_weight_wet_substrateremoved)) +
      (ifelse(is.na(prelim.data$Flush4Fruitbody_weight_wet), 0, prelim.data$Flush4Fruitbody_weight_wet) -
       ifelse(is.na(prelim.data$Flush4Fruitbody_weight_wet_substrateremoved), 0, prelim.data$Flush4Fruitbody_weight_wet_substrateremoved)) +
      (ifelse(is.na(prelim.data$AsexualGrowthWeightWet), 0, prelim.data$AsexualGrowthWeightWet) -
       ifelse(is.na(prelim.data$AsexualGrowthWeightWet_substrateremoved), 0, prelim.data$AsexualGrowthWeightWet_substrateremoved)) +
      (ifelse(is.na(prelim.data$ImmatureHarvestWetWEight), 0, prelim.data$ImmatureHarvestWetWEight) -
       ifelse(is.na(prelim.data$ImmatureHarvestWetWEight_substrateremoved), 0, prelim.data$ImmatureHarvestWetWEight_substrateremoved))
  ) * (prelim.data$SubstrateEndWeightDry_raw/prelim.data$SubstrateEndWeightWet_raw)

## Substrate mass and substrate mass loss
prelim.data$SubstrateMassLossPercentage <- (prelim.data$SubstrateMassLoss)/prelim.data$SubstrateMassInitial
```


## Calculating Spore Load (sporeprint weight ratio)
```{r}
# Formula is Sporeprint dry weight / (SubsampleWetWeight - SubsampleWetWeight * Water Content)
prelim.data$SporeLoad <- prelim.data$Sporeprint_weight_dry/(prelim.data$Sporeprint_subsampl_weight- prelim.data$Sporeprint_subsampl_weight*prelim.data$Flush1Fruitbody_water_content)

# Removing outlier DH1153 (error in data entry possibly)
prelim.data$SporeLoad[prelim.data$SporeLoad > 1.0] <- NA

# Changing zeroes and negative values to zero
#prelim.data$SporeLoad[prelim.data$SporeLoad < 0.0001] <- NA

# Adding the only true zeroes I've recorded for ICMP11565 which no spores were observed
prelim.data$SporeLoad[prelim.data$Location_code == "ICMP11565"] <- 0
```

## Import image metrics: colour and fruitbody size from other R markdown file
Ex1_colordistance_import.Rmd

```{r}
# Objects are: Flush1_Fruitbody_size and RGB_values_Flush1. Both have an Experiment1Positition column that I can use to match data
Flush1_Fruitbody_size <- readRDS(file = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/R project/Robjects/Flush1_Fruitbody_size.rds")
RGB_values_Flush1 <- readRDS(file = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/R project/Robjects/RGB_values_Flush1.rds")

prelim.data <- left_join(prelim.data, Flush1_Fruitbody_size, by = "Experiment1Positition")
prelim.data <- left_join(prelim.data, RGB_values_Flush1, by = "Experiment1Positition")
```


## loading hardness measurements
Hardness measurements from texture analyser are all stored in a specific cell in individual xlsx files

```{r}
setwd("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/raw data/texture measurements/")
hardnessList <- list.files("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/raw data/texture measurements/")

hardnessData = lapply(hardnessList, function(i){
    x = read.xlsx(i, sheetIndex=1, sheetName=NULL, startRow=10,
        endRow=11, as.data.frame=TRUE, header=T)
    # Get the columns you want: 13 (M)
    x = x[, 11]
    # You may want to add a column to say which file they're from
    x$Experiment1Positition = basename(i)
    # Return your data
    x
})

hardnessDataF = do.call("rbind.data.frame", hardnessData)

hardnessDataF$Experiment1Positition <- sub(".xls", "", hardnessDataF$Experiment1Positition)

colnames(hardnessDataF)[1] <- 'Flush1_Hardness'
```

## Adding hardness measurements into the rest of my datasheet
```{r}
merged_prelimData <- full_join(prelim.data, hardnessDataF, by = "Experiment1Positition")
```


## Remove outliers
```{r}
# Dja dry and tiny - "C1-22  DH814 Dja" - driest fruitbody - "C2-21 DH803 Dja" - tiny fruitbody

# Identify rows where Experiment1Position is "C1-22" or "C2-21"
rows <- merged_prelimData$Experiment1Positition %in% c("C1-22", "C2-21")

# Replace the corresponding values in Flush1_Hardness and Flush1Fruitbody_water_content with NAs
merged_prelimData$Flush1_Hardness[rows] <- NA
merged_prelimData$Flush1Fruitbody_water_content[rows] <- NA

```


# global variables and vectors

Global color palette for my 5 species and appropriate levels
```{r}
SpeciesColors = c("#FB9A99", "#E69F00", "#56B4E9", "#009E73", "#D55E00", "grey")
four_colors = c("#FB9A99", "#E69F00", "#56B4E9", "#009E73")
ColorLevels = c("Dja", "Pars", "Pul", "Ppo", "Aus", "Control")
SpeciesColorsPul = c("#FB9A99", "#E69F00", "#56B4E9", "#4C7113", "#009E73", "#D55E00", "grey")
ColorLevelsPul = c("Dja", "Pars", "Pul_NZ", "Pul_intl", "Ppo", "Aus", "Control")
merged_prelimData$Species.Code <- factor(merged_prelimData$Species.Code, levels = ColorLevels)
filtered_Data$Species.Code <- factor(filtered_Data$Species.Code, levels = ColorLevels)
names(SpeciesColors) <- ColorLevels
filtered_Data$Species <- factor(filtered_Data$Species, levels = ColorLevelsPul)
names(SpeciesColorsPul) <- ColorLevelsPul
specieslabels = c("Dja" = "P. djamor", "Pars" = "P. parsonsiae", "Pul_NZ" = "P. pulmonarius NZ", "Pul_intl" = "P. pulmonarius intl.", "Ppo" = "P. purpureo-olivaceus", "Aus" = "P. australis")
specieslabelsRepr = c("Dja" = "P. djamor", "Pars" = "P. parsonsiae", "Pul_NZ" = "P. pulmonarius NZ", "Pul_intl" = "P. pulmonarius intl.", "Aus" = "P. australis")
```

#Global trait list
```{r}
includeTraits <- c("SubstratePointsColonizedT1", "SubstratePointsColonizedT2", "SubstrateMassLossPercentage",
                   "Flush1Time_to_Fruiting", "Flush1Time_to_Mature", "Flush1Fruitbody_number", 
                   "Flush1_Fruitbody_size", "Flush1Fruitbody_total_dryweight", "Flush1_Hardness", 
                   "Flush1Fruitbody_water_content", "TotalBiomassProducedDryWeight", "Cap_r", "Cap_g", 
                   "Cap_b", "Gill_r", "Gill_g", "Gill_b")

traitLabels <- c("Substrate colonised (day 8)", "Substrate colonised (day 16)", "Substrate mass loss (%)",
                 "Time to fruit", "Time to mature", "Number of fruitbodies", "Fruitbody size",
                 "Fruitbody weight", "Fruitbody hardness", "Fruitbody water content", "Total biomass produced",
                 "Cap colour - Reds", "Cap colour - Greens", "Cap colour - Blues", "Gill colour - Reds",
                 "Gill colour - Greens", "Gill colour - Blues")

traitLabelsNoUnits <- c("Substrate colonised (day 8)", "Substrate colonised (day 16)", "Substrate mass loss",
                        "Time to fruit", "Time to mature", "Number of fruitbodies", "Fruitbody size",
                        "Fruitbody weight", "Fruitbody hardness", "Fruitbody water content", "Total biomass produced", "Cap colour - Reds", "Cap colour - Greens", "Cap colour - Blues", "Gill colour - Reds", "Gill colour - Greens", "Gill colour - Blues")

includeReprTraits_Named <- c("Time to fruit (d)", "Time to mature (d)", "Number of Sporocarps","Cap colour - Reds (%)", "Cap colour - Greens (%)", "Cap colour - Blues (%)", "Gill colour - Reds (%)", "Gill colour - Greens (%)", "Gill colour - Blues (%)", "Sporocarp hardness (g)", "Sporocarp water content (%)",  "Sporocarp size (mm2)", "Sporocarp weight (g)", "Total biomass produced (g)")

includeColourtraits <- c("Cap_r", "Cap_g", "Cap_b", "Gill_r", "Gill_g", "Gill_b")

includeReprTraits <- c("Flush1Time_to_Fruiting", "Flush1Time_to_Mature", "Flush1Fruitbody_number","Cap_r", "Cap_g", "Cap_b", "Gill_r", "Gill_g", "Gill_b", "Flush1_Hardness", "Flush1Fruitbody_water_content",  "Flush1_Fruitbody_size", "Flush1Fruitbody_total_dryweight", "TotalBiomassProducedDryWeight")



includeVegTraits <- c("SubstrateMassLossPercentage", "SubstratePointsColonizedT1", "SubstratePointsColonizedT2")
includeVegTraits_Named <- c("Substrate mass loss (%)", "Substrate colonised (day 8)", "Substrate colonised (day 16)")

includeGroups <- c("Experiment1Positition", "Sample.source", "Species.Code", "Species")

includeGroupsandTraits <- c("Experiment1Positition", "Sample.source", "Species.Code", includeTraits)
```

# Calculating number of strains that fruited all 3 replicates
```{r}
#Subset data to only focus on Dja, Pars and Pul
repstrains <- c("Aus", "Ppo", "Control")
calculation_data <- merged_prelimData[!(merged_prelimData$Species.Code %in% c("Control", "Ppo", "Aus")), ]

# Group by Sample.Source and count the number of non-NA Flush1PinningDate entries for each strain
count_data <- calculation_data %>%
  group_by(Sample.source) %>%
  summarise(replicate_count = sum(!is.na(Flush1PinningDate)), .groups = 'drop')

# Filter to keep only those strains that have 3 replicates
successful_strains <- count_data %>% filter(replicate_count == 3)

# Count the number of unique strains that have successfully fruited in all three replicates
num_successful_strains <- nrow(successful_strains)

# Print the result
print(num_successful_strains)

# Calculate mean and standard deviation of the replicate counts
mean_replicates <- mean(count_data$replicate_count)
std_replicates <- sd(count_data$replicate_count)

# Print the results
cat("Number of strains where all 3 replicates fruited", num_successful_strains, "\n")
cat("Proportion of strains where all 3 replicates fruited", num_successful_strains/45, "\n")
cat("Mean count of replicates that fruited:", mean_replicates, "\n")
cat("Standard deviation of replicates that fruited:", std_replicates, "\n")
```


# Filtering my source data to only include relevant data for my analysis

Create one dataframe per trait for filtering in the next step

The following code creates one new dataframe for each trait from my global traits list object includeTraits. All the trait specific dataframes then get combined into a new master dataframe called filtered_Data.
```{r}
for (i in includeTraits) {
  nmissingtable <- table(merged_prelimData$Sample.source, is.na(merged_prelimData[i])) # check which observations have no records, for a given trait
  includeStrain <- row.names(nmissingtable)[nmissingtable[,1] > 2] # create list of strains that have 3 or more records
  tempPlottingData <- merged_prelimData[merged_prelimData$Sample.source %in% includeStrain,] # subsetting from original dataframe
  tempPlottingData <- tempPlottingData[, names(tempPlottingData) %in% c("Experiment1Positition", i)] # selecting only position ID and trait data
  assign(paste0(i), tempPlottingData) # create a new name for the output dataframe, which has two variables: position ID and data for the i trait
}

```

Creating a new dataframe with the filtered data temp trait objects from above #baseRftw!
```{r}
rm(filtered_Data)
filtered_Data <- merged_prelimData[, c(1:9)] # create new dataframe with just Experiment1Position ID until column 9 ("Species")
#colnames(filtered_Data) <- c("Experiment1Positition")

for (i in includeTraits) {
    tempTrait <- get(i) # find object with same name as trait
filtered_Data[i] <- tempTrait[, c(2)][match(filtered_Data$Experiment1Positition, tempTrait$Experiment1Positition)] # match() is used to look for values in first dataframe, compare them to values in second dataframe, and then populate data in chosen column 
}

## Adding shelf position to data
positionA <- as.factor(substr(filtered_Data$Experiment1Positition, 1, 1)) # shelf position (1-24)
positionB <- as.factor(substr(filtered_Data$Experiment1Positition, 2, 2)) # shelf position (1-24)
filtered_Data$shelfxlevel <- paste0(positionA,"", positionB)
filtered_Data$shelf <- paste0(positionA)
filtered_Data$level <- paste0(positionB)
```

## Adding shelf position to data

```{r}
#filtered_Data$shelf <- as.factor(substr(filtered_Data$Experiment1Positition, 1, 1)) # shelf ID (A-D)
#filtered_Data$shelflevel <- as.factor(substr(filtered_Data$Experiment1Positition, 2, 2)) # shelf level (1-5)
#positionA <- as.factor(substr(filtered_Data$Experiment1Positition, 4, 4)) # shelf position (1-24)
#positionB <- as.factor(substr(filtered_Data$Experiment1Positition, 5, 5)) # shelf position (1-24)
#filtered_Data$shelfposition <- paste0(positionA,"", positionB)

positionA <- as.factor(substr(filtered_Data$Experiment1Positition, 1, 1)) # shelf position (1-24)
positionB <- as.factor(substr(filtered_Data$Experiment1Positition, 2, 2)) # shelf position (1-24)
filtered_Data$shelfxlevel <- paste0(positionA,"", positionB)
filtered_Data$shelf <- paste0(positionA)
filtered_Data$level <- paste0(positionB)
#filtered_Data$shelfxlevel <- gsub("^(.{1})5", "\\14", filtered_Data$shelfxlevel) #merge shelf bottom layer with one above (4 and 5) because level 5 only has 3 observations

```


```{r}
#Remove objects with trait names again
rm(Flush1_Hardness)
rm(Flush1Time_to_Mature)
rm(Flush1Time_to_Fruiting)
rm(SubstrateMassLossPercentage)
rm(Flush1Fruitbody_total_dryweight)
rm(Flush1Fruitbody_water_content)
```

# Calculating number of strains that fruited all 3 replicates
```{r}
#Subset data to only focus on Dja, Pars and Pul
repstrains <- c("Aus", "Ppo", "Control")
calculation_data <- filtered_Data[!(filtered_Data$Species.Code %in% c("Control", "Ppo", "Aus")), ]

# Group by Sample.Source and count the number of non-NA Flush1PinningDate entries for each strain
count_data <- calculation_data %>%
  group_by(Sample.source) %>%
  summarise(replicate_count = sum(!is.na(Flush1Time_to_Mature)), .groups = 'drop')

# Filter to keep only those strains that have 3 replicates
successful_strains <- count_data %>% filter(replicate_count == 3)

# Count the number of unique strains that have successfully fruited in all three replicates
num_successful_strains <- nrow(successful_strains)

# Print the result
print(num_successful_strains)

# Calculate mean and standard deviation of the replicate counts
mean_replicates <- mean(count_data$replicate_count)
std_replicates <- sd(count_data$replicate_count)

# Print the results
cat("Number of strains where all 3 replicates fruited", num_successful_strains, "\n")
cat("Proportion of strains where all 3 replicates fruited", num_successful_strains/45, "\n")
cat("Mean count of replicates that fruited:", mean_replicates, "\n")
cat("Standard deviation of replicates that fruited:", std_replicates, "\n")
```

# Individual trait results

## Distance matrix (Euclidean) for testing hypothesis 1 (within-species variation in reproductive vs vegetative traits)
```{r}
# Remove identifiers from data, using my transformed data (normalised data for distance)
# exclude Australis
filtered_Data_NoAus <- filtered_Data[!filtered_Data$Species.Code == "Aus",]
#filtered_Data_NoAus <- filtered_Data
filtered_Data_NoAus <- filtered_Data_NoAus[!filtered_Data_NoAus$Species.Code == "Control",]
filtered_Data_NoAus <- filtered_Data_NoAus[!filtered_Data_NoAus$Species.Code == "Ppo",]
filtered_Data_NoAus <- filtered_Data_NoAus[order(filtered_Data_NoAus$Species.Code),] # order by Species
data_for_distance <- filtered_Data_NoAus
data_for_distance$Species.Code <- factor(data_for_distance$Species.Code, levels = ColorLevels)
data_for_distance$Species <- factor(data_for_distance$Species, levels = ColorLevelsPul)

# Filter for cultivated/wild testing
data_for_hyp2 <- data_for_distance[!data_for_distance$Species == "Pul_intl",]
data_for_hyp2 <- data_for_hyp2[!data_for_hyp2$Species == "Dja",]
data_for_hyp2 <- data_for_hyp2[!apply(is.na(data_for_hyp2[includeTraits]), 1, any), ] # remove any rows with NAs
data_scaled_hyp2 <- decostand(data_for_hyp2[includeTraits], method = "standardize")

## Scale my data using decostand()
data_scaled <- decostand(data_for_distance[includeTraits], method = "standardize")
data_scaled_hyp2 <- decostand(data_for_hyp2[includeTraits], method = "standardize")
# log transform first
#data_scaled <- decostand(log(data_for_distance[includeTraits]), method = "standardize")

# create labels dataframe
sample_labels <- as.data.frame(data_for_distance$Species.Code)
sample_labels$Strains <- data_for_distance$Sample.source
sample_labels <- rename(sample_labels, "Species" = "data_for_distance$Species.Code")

# Categorizing traits by vegetative/reproductive
vegetative_traits <- c("SubstratePointsColonizedT1", "SubstratePointsColonizedT2", "SubstrateMassLossPercentage")
reproductive_traits <- c("Flush1Time_to_Fruiting", "Flush1Time_to_Mature", "Flush1Fruitbody_total_dryweight", "TotalBiomassProducedDryWeight", "Flush1_Hardness", "Flush1_Fruitbody_size", "Cap_r", "Cap_g", "Cap_b", "Gill_r", "Gill_g", "Gill_b", "Flush1Fruitbody_number", "Flush1Fruitbody_water_content")

# Split data into vegetative and reproductive
vegetative_results <- data_scaled[, vegetative_traits]
vegetative_results <- vegetative_results[!apply(is.na(vegetative_results[vegetative_traits]), 1, any), ] # remove any rows with NAs
#vegetative_results$SubstratePointsColonizedT1 <- as.numeric(vegetative_results$SubstratePointsColonizedT1)
#vegetative_results$SubstratePointsColonizedT2 <- as.numeric(vegetative_results$SubstratePointsColonizedT2)
reproductive_results <- data_scaled[, reproductive_traits]
reproductive_results <- reproductive_results[!apply(is.na(reproductive_results[reproductive_traits]), 1, any), ] # remove any rows with NAs

# remove any rows with NAs
data_scaled_noNA <- data_scaled[!apply(is.na(data_scaled[includeTraits]), 1, any), ] # remove any rows with NAs

# remove any rows with NAs
data_for_alltraits_noNA <- data_for_distance[!apply(is.na(data_for_distance[includeTraits]), 1, any), ] # remove any rows with NAs

# Remove NAs from all traits data
data_scaled_NArm <- data_scaled[!apply(is.na(data_scaled[includeTraits]), 1, any), ] # remove any rows with NAs

#rename traits into full descriptions
for(i in 1:length(includeTraits)){
  if(includeTraits[i] %in% colnames(vegetative_results)){
    colnames(vegetative_results)[colnames(vegetative_results) == includeTraits[i]] <- traitLabelsNoUnits[i]
  }
}

for(i in 1:length(includeTraits)){
  if(includeTraits[i] %in% colnames(reproductive_results)){
    colnames(reproductive_results)[colnames(reproductive_results) == includeTraits[i]] <- traitLabelsNoUnits[i]
  }
}

# subsetting original data for only rows with vegetative traits data
data_for_vegetative <- data_for_distance[!apply(is.na(data_for_distance[vegetative_traits]), 1, any), ] # remove any rows with NAs
data_for_vegetative$Species.Code <- factor(data_for_vegetative$Species.Code, levels = ColorLevels)
#data_for_vegetative <- data_for_vegetative %>% #rename species acronyms into full species names
#  mutate(Species.Code = as.character(Species.Code)) %>%
#  mutate(Species.Code = ifelse(Species.Code %in% names(specieslabels), specieslabels[Species.Code], Species.Code)) %>%
#  mutate(Species.Code = as.factor(Species.Code))
#colnames(data_for_vegetative)[colnames(data_for_vegetative) %in% includeTraits] <- traitLabels # rename traits

# subsetting original data for only rows with reproductive traits data
data_for_reproductive <- data_for_distance[!apply(is.na(data_for_distance[reproductive_traits]), 1, any), ] # remove any rows with NAs
data_for_reproductive$Species.Code <- factor(data_for_reproductive$Species.Code, levels = ColorLevels)

data_for_reproductive <- data_for_reproductive %>%
  mutate(Species.Code = as.character(Species.Code)) %>%
  mutate(Species.Code = ifelse(Species.Code %in% names(specieslabels), specieslabels[Species.Code], Species.Code)) %>%
  mutate(Species.Code = as.factor(Species.Code))

# calculating Gower's distance matrix for each group
vegetative_daisy <- daisy(vegetative_results, metric = "euclidean")
vegetative_daisy_matrix <- as.matrix(vegetative_daisy)
reproductive_daisy <- daisy(reproductive_results, metric = "euclidean")
reproductive_daisy_nowater <- daisy(reproductive_results[,-c(14)], metric = "euclidean")
reproductive_daisy_matrix <- as.matrix(reproductive_daisy)


# PERMANOVA 
vegetative_permanova <- adonis2(vegetative_daisy_matrix ~ Species/Sample.source, data = data_for_vegetative)
print(vegetative_permanova)
reproductive_permanova <- adonis2(reproductive_daisy_matrix ~ Species/Sample.source, data = data_for_reproductive)
print(reproductive_permanova)
```



## Color coding of strains
Create color palette for all strains per species

```{r}
# Sort dataframe by Species
virdis_data <- filtered_Data %>%
  arrange(Sample.source) %>%
  #filter(!(Species.Code %in% c("Control", NA)))
  filter(!(Species.Code %in% c("Control", "Aus", "Ppo", NA)))

# Create unique color for strains
num_colors <- length(unique(virdis_data$Sample.source))
strain_colors <- viridis::viridis_pal(begin = 0.25, end = 0.9, option = "plasma")(num_colors)

# Step 2: Create a named vector of colors for Sample.source
unique_sources <- unique(virdis_data$Sample.source)
names(strain_colors) <- unique_sources

# Create unique color for traits
#num_colors <- length(unique(trait_ranges$Trait))
#trait_colors <- viridis::viridis_pal(begin = 0.05, end = 0.9, option = "turbo")(num_colors)

# Step 2: Create a named vector of colors for Sample.source
#unique_sources <- unique(trait_ranges$Trait)
#names(trait_colors) <- unique_sources



# Filter out 'Control' and NA from Species
virdis_data <- filtered_Data %>%
  filter(!(Species %in% c("Control", NA))) %>%
  arrange(Species, Sample.source)

# Get unique species and color options
unique_species <- unique(virdis_data$Species)
color_options <- c("magma", "cividis", "mako", "viridis", "inferno", "plasma")

# Initialize an empty list to store color vectors
species_color_vectors <- list()

# Loop through each species to create a color vector for its Sample.source
for (i in seq_along(unique_species)) {
  species <- unique_species[i]
  
  # Subset data for the species
  species_data <- filter(virdis_data, Species == species)
  
  # Determine the number of unique Sample.source within this species
  num_colors <- length(unique(species_data$Sample.source))
  
  # Choose a color option, cycling through color_options if necessary
  color_option <- color_options[(i - 1) %% length(color_options) + 1]
  
  # Generate color palette
  strain_colors <- viridis::viridis_pal(begin = 0.35, end = 0.85, option = color_option)(num_colors)
  
  # Create a named vector of colors for Sample.source within this species
  unique_sources <- unique(species_data$Sample.source)
  names(strain_colors) <- unique_sources
  
  # Store the color vector in the list, named by the species
  species_color_vectors[[species]] <- strain_colors
}

# Combine all species' color vectors into a single named vector
all_strain_colors <- unlist(species_color_vectors, use.names = TRUE)
```

## Raincloud plot ggrain - test on a single trait
```{r}
plot_test <- ggplot(filtered_Data[!(filtered_Data$Species %in% c("Control", "Ppo", NA)), ], aes(x=Species, y=Flush1Time_to_Mature)) +
  geom_rain(alpha = .5, point.args = list(alpha = 0), 
            boxplot.args = list(alpha = 0, color = "white", outlier.shape = NA), 
            violin.args = list(aes(fill = Species), alpha = 0.8),
            violin.args.pos = list(
              side = "r",
              width = 0.7, position = position_nudge(x = 0.26))) +
  geom_boxplot(aes(group=Sample.source, color = Sample.source), outlier.shape = NA, alpha = .4, width = .5, position = position_dodge(width = 0.45)) + 
  geom_point(aes(color = Sample.source), position = position_dodge(width = 0.45), size = .8) +
  scale_fill_manual(values = SpeciesColorsPul) +
  scale_color_manual(values = all_strain_colors) +
  guides(color = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  #ggplot2::ylim(0, 10)
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 1))

print(plot_test, dpi = 300)
```


## Facetplots violin with raincloud - vegetative traits
```{r}
# Assuming filtered_Data is your original dataframe
vegetative_plots_data <- filtered_Data[!apply(is.na(filtered_Data[vegetative_traits]), 1, any), ] # remove any rows with NAs

# Calculate the number of unique Sample.source for each Species
species_counts_veg <- aggregate(Sample.source ~ Species, data = vegetative_plots_data, FUN = function(x) length(unique(as.character(x))))
names(species_counts_veg)[2] <- "n"
species_counts_veg$Species_label <- paste(specieslabels, "\n ( n =", species_counts_veg$n,")")


# Reshape data to long format
long_Data <- filtered_Data %>%
  pivot_longer(cols = includeVegTraits, names_to = "Trait", values_to = "TraitValue") %>%
  #pivot_longer(cols = includeTraitsWOcolour, names_to = "Trait", values_to = "TraitValue") %>%
  filter(!(Species.Code %in% c("Control", NA))) 
  #filter(!(Species %in% c("Control", "Aus", "Ppo", NA))) 

# Rename traits in long_Data using includeVegTraits and includeVegTraits_Named
trait_mapping <- setNames(includeVegTraits_Named, includeVegTraits)
long_Data <- long_Data %>%
  mutate(Trait = replace(Trait, Trait %in% names(trait_mapping), trait_mapping[Trait]))

# Create a list to hold individual plots
grob_list <- list()

# Create individual plots for each trait
traits <- unique(long_Data$Trait)
for (trait in traits) {
  plot <- ggplot(subset(long_Data, Trait == trait), aes(x = Species, y = TraitValue)) +
    geom_rain(alpha = .5, adjust = 0.3, point.args = list(alpha = 0),
              boxplot.args = list(alpha = 0, color = "white", outlier.shape = NA),
              violin.args = list(aes(fill = Species), size = 0.1, alpha = 0.66, adjust = 0.77),
              violin.args.pos = list(
                side = "r",
                width = 0.7, position = position_nudge(x = 0.22))) +
    geom_line(aes(group = Sample.source, color = Sample.source), position = position_dodge(width = 0.35), alpha = 0.5) +
    geom_point(aes(color = Sample.source), position = position_dodge(width = 0.35), size = 0.8, alpha = 0.5) +
    scale_fill_manual(values = SpeciesColorsPul) +
    scale_color_manual(values = all_strain_colors) +
    scale_x_discrete(labels = species_counts_veg$Species_label) +
    guides(color = FALSE) +
    theme_minimal() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 9)) +
    labs(y = trait)
  
  grob_list[[trait]] <- plot
}

# Arrange the plots side by side using grid.arrange
grid.arrange(grobs = grob_list, ncol = 3)

# Arrange the plots side by side using grid.arrange and save as PNG
png(filename = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/finalplots/plot_facet_rain_veg_2Pul.png", 
    width = 9, height = 3, units = "in", res = 300, bg = "white")
grid.arrange(grobs = grob_list, ncol = 3)
dev.off()

```
#### Raincloud with CLD
```{r}
# Assuming filtered_Data is your original dataframe
vegetative_plots_data <- filtered_Data[!apply(is.na(filtered_Data[vegetative_traits]), 1, any), ] # remove any rows with NAs

# Calculate the number of unique Sample.source for each Species
species_counts_veg <- aggregate(Sample.source ~ Species, data = vegetative_plots_data, FUN = function(x) length(unique(as.character(x))))
names(species_counts_veg)[2] <- "n"
species_counts_veg$Species_label <- paste(specieslabels, "\n ( n =", species_counts_veg$n,")")

# Reshape data to long format
long_Data <- filtered_Data %>%
  pivot_longer(cols = includeVegTraits, names_to = "Trait", values_to = "TraitValue") %>%
  #pivot_longer(cols = includeTraitsWOcolour, names_to = "Trait", values_to = "TraitValue") %>%
  filter(!(Species %in% c("Control", NA))) 
  #filter(!(Species %in% c("Control", "Aus", "Ppo", NA))) 

# Rename traits in long_Data using includeVegTraits and includeVegTraits_Named
trait_mapping <- setNames(includeVegTraits_Named, includeVegTraits)
long_Data <- long_Data %>%
  mutate(Trait = factor(replace(Trait, Trait %in% names(trait_mapping), trait_mapping[Trait]), levels = includeVegTraits_Named))

# Perform the statistical test (ANOVA and Tukey HSD in this case)
aov_results <- long_Data %>%
  group_by(Trait) %>%
  summarise(tukey = list(TukeyHSD(aov(TraitValue ~ Species))))

# Function to calculate compact letter displays for each trait
calculate_cld <- function(tukey_result) {
  tukey_diff <- tukey_result$Species[, "diff"]
  tukey_p_adj <- tukey_result$Species[, "p adj"]
  names(tukey_diff) <- rownames(tukey_result$Species)
  names(tukey_p_adj) <- rownames(tukey_result$Species)
  
  comparison_matrix <- matrix(NA, nrow = length(unique(unlist(strsplit(names(tukey_diff), "-")))), ncol = length(unique(unlist(strsplit(names(tukey_diff), "-")))))
  rownames(comparison_matrix) <- unique(unlist(strsplit(names(tukey_diff), "-")))
  colnames(comparison_matrix) <- unique(unlist(strsplit(names(tukey_diff), "-")))
  
  for (i in seq_along(tukey_diff)) {
    comparison <- strsplit(names(tukey_diff)[i], "-")[[1]]
    comparison_matrix[comparison[1], comparison[2]] <- tukey_p_adj[i]
    comparison_matrix[comparison[2], comparison[1]] <- tukey_p_adj[i]
  }
  
  cld <- multcompLetters(comparison_matrix, threshold = 0.05)$Letters
  cld <- as.data.frame(cld)
  cld$Species <- rownames(cld)
  rownames(cld) <- NULL
  return(cld)
}

# Initialize an empty dataframe to store all CLDs
all_clds <- data.frame()

# Loop through each trait and calculate CLDs
for (i in 1:nrow(aov_results)) {
  cld <- calculate_cld(aov_results$tukey[[i]])
  cld$Trait <- aov_results$Trait[i]
  all_clds <- rbind(all_clds, cld)
}

# Merge the CLD data with the main data
long_Data <- long_Data %>%
  left_join(all_clds, by = c("Trait", "Species"))

long_Data$Species <- factor(long_Data$Species, levels = ColorLevelsPul)

# Create a list to hold individual plots
grob_list <- list()

# Create individual plots for each trait
traits <- unique(long_Data$Trait)
for (i in seq_along(traits)) {
  trait <- traits[i]
  trait_data <- subset(long_Data, Trait == trait)
  lowest_y_value <- min(trait_data$TraitValue, na.rm = TRUE)
  y_position <- lowest_y_value * 0.9
  
  plot <- ggplot(subset(long_Data, Trait == trait), aes(x = Species, y = TraitValue)) +
    geom_rain(alpha = .5, adjust = 0.3, point.args = list(alpha = 0),
              boxplot.args = list(alpha = 0, color = "white", outlier.shape = NA),
              violin.args = list(aes(fill = Species), size = 0.1, alpha = 0.66, adjust = 0.77),
              violin.args.pos = list(
                side = "r",
                width = 0.7, position = position_nudge(x = 0.22))) +
    geom_line(aes(group = Sample.source, color = Sample.source), position = position_dodge(width = 0.35), alpha = 0.5) +
    geom_point(aes(color = Sample.source), position = position_dodge(width = 0.35), size = 0.8, alpha = 0.5) +
    geom_text(aes(label = cld, y = y_position), size = 3) + # Add CLD labels 
    scale_fill_manual(values = SpeciesColorsPul) +
    scale_color_manual(values = all_strain_colors) +
    scale_x_discrete(labels = species_counts_veg$Species_label) +
    guides(color = FALSE) +
    theme_minimal() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 8)) +
    labs(y = trait)
  
  # Add an uppercase letter annotation to each plot
  plot <- plot + annotate("text", x = -Inf, y = Inf, label = LETTERS[i], hjust = -0.5, vjust = 1.5, size = 6, fontface = "bold")
  
  grob_list[[i]] <- plot
  #grob_list[[trait]] <- plot
}

# Arrange the plots side by side using grid.arrange
grid.arrange(grobs = grob_list, ncol = 3)

# Arrange the plots side by side using grid.arrange and save as PNG
png(filename = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/finalplots/plot_facet_rain_veg_2Pul_CLD.png", 
    width = 9, height = 3, units = "in", res = 300, bg = "white")
grid.arrange(grobs = grob_list, ncol = 3)
dev.off()
```


## Facetplots violin with raincloud - reproductive traits
```{r}
# Assuming filtered_Data is your original dataframe
reproductive_plots_data <- filtered_Data[!apply(is.na(filtered_Data[includeReprTraits]), 1, any), ] # remove any rows with NAs

# Calculate the number of unique Sample.source for each Species
species_counts_repr <- aggregate(Sample.source ~ Species, data = reproductive_plots_data, FUN = function(x) length(unique(as.character(x))))
names(species_counts_repr)[2] <- "n"
species_counts_repr$Species_label <- paste(specieslabelsRepr, "\n ( n =", species_counts_repr$n,")")

# Reshape data to long format
long_Data <- filtered_Data %>%
  pivot_longer(cols = includeReprTraits, names_to = "Trait", values_to = "TraitValue") %>%
  #pivot_longer(cols = includeTraitsWOcolour, names_to = "Trait", values_to = "TraitValue") %>%
  #filter(!(Species %in% c("Control", NA))) # Applying your existing filter
  filter(!(Species %in% c("Control", "Aus", "Ppo", NA))) # Applying your existing filter

# Update plot code to use the reshaped data
plot_facet_repr <- ggplot(long_Data, aes(x = Species, y = TraitValue)) +
  geom_rain(alpha = .5, adjust = 0.3, point.args = list(alpha = 0), 
            boxplot.args = list(alpha = 0, color = "white", outlier.shape = NA), 
            violin.args = list(aes(fill = Species), size = 0.1, alpha = 0.66, adjust = 0.77),
            violin.args.pos = list(
              side = "r",
              width = 0.7, position = position_nudge(x = 0.22))) +
  #geom_boxplot(aes(group=Sample.source, color = Sample.source), outlier.shape = NA, alpha = .7, width = .5) +
  geom_line(aes(group = Sample.source, color = Sample.source), position = position_dodge(width = 0.35), alpha = 0.5) +
  #geom_point(aes(color = Sample.source), position = position_jitter(width = .15), size = .8) +
  geom_point(aes(color = Sample.source), position = position_dodge(width = 0.35), size = .8, alpha = 0.5) +
  scale_fill_manual(values = SpeciesColorsPul) +
  scale_color_manual(values = all_strain_colors) +
  scale_x_discrete(labels = species_counts_repr$Species_label) +
  facet_wrap(~Trait, scales = "free", shrink = TRUE, ncol = 3) + # Set max of 3 rows of graphs
  guides(color = FALSE) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))  

# Print the plot
print(plot_facet_repr)

ggsave(filename = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.facet_rain_repr_2Pul.png", plot = plot_facet_repr, width = 10, height = 12, dpi = 300, bg = "white")
#ggsave(filename = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.facet_rain_scaled.png", plot = plot_facet_repr, width = 10, height = 12, dpi = 300, bg = "white")
```
#### Raincloud with CLD grob
```{r}
# Assuming filtered_Data is your original dataframe
reproductive_plots_data <- filtered_Data[!apply(is.na(filtered_Data[reproductive_traits]), 1, any), ] # remove any rows with NAs

# Calculate the number of unique Sample.source for each Species
species_counts_repr <- aggregate(Sample.source ~ Species, data = reproductive_plots_data, FUN = function(x) length(unique(as.character(x))))
names(species_counts_repr)[2] <- "n"
species_counts_repr$Species_label <- paste(specieslabelsRepr, "\n ( n =", species_counts_repr$n,")")

# Reshape data to long format
long_Data <- filtered_Data %>%
  pivot_longer(cols = includeReprTraits, names_to = "Trait", values_to = "TraitValue") %>%
  #pivot_longer(cols = includeTraitsWOcolour, names_to = "Trait", values_to = "TraitValue") %>%
  #filter(!(Species %in% c("Control", NA))) # Applying your existing filter
  filter(!(Species %in% c("Control", "Aus", "Ppo", NA))) # Applying your existing filter

# Rename traits in long_Data using includeVegTraits and includeVegTraits_Named
trait_mapping <- setNames(includeReprTraits_Named, includeReprTraits)
long_Data <- long_Data %>%
  mutate(Trait = replace(Trait, Trait %in% names(trait_mapping), trait_mapping[Trait]))

# Perform the statistical test (ANOVA and Tukey HSD in this case)
aov_results <- long_Data %>%
  group_by(Trait) %>%
  summarise(tukey = list(TukeyHSD(aov(TraitValue ~ Species))))

# Function to calculate compact letter displays for each trait
calculate_cld <- function(tukey_result) {
  tukey_diff <- tukey_result$Species[, "diff"]
  tukey_p_adj <- tukey_result$Species[, "p adj"]
  names(tukey_diff) <- rownames(tukey_result$Species)
  names(tukey_p_adj) <- rownames(tukey_result$Species)
  
  comparison_matrix <- matrix(NA, nrow = length(unique(unlist(strsplit(names(tukey_diff), "-")))), ncol = length(unique(unlist(strsplit(names(tukey_diff), "-")))))
  rownames(comparison_matrix) <- unique(unlist(strsplit(names(tukey_diff), "-")))
  colnames(comparison_matrix) <- unique(unlist(strsplit(names(tukey_diff), "-")))
  
  for (i in seq_along(tukey_diff)) {
    comparison <- strsplit(names(tukey_diff)[i], "-")[[1]]
    comparison_matrix[comparison[1], comparison[2]] <- tukey_p_adj[i]
    comparison_matrix[comparison[2], comparison[1]] <- tukey_p_adj[i]
  }
  
  cld <- multcompLetters(comparison_matrix, threshold = 0.05)$Letters
  cld <- as.data.frame(cld)
  cld$Species <- rownames(cld)
  rownames(cld) <- NULL
  return(cld)
}

# Initialize an empty dataframe to store all CLDs
all_clds <- data.frame()

# Loop through each trait and calculate CLDs
for (i in 1:nrow(aov_results)) {
  cld <- calculate_cld(aov_results$tukey[[i]])
  cld$Trait <- aov_results$Trait[i]
  all_clds <- rbind(all_clds, cld)
}

# Merge the CLD data with the main data
long_Data <- long_Data %>%
  left_join(all_clds, by = c("Trait", "Species"))

long_Data$Species <- factor(long_Data$Species, levels = ColorLevelsPul)

# Create a list to hold individual plots
grob_list <- list()

# Create individual plots for each trait
traits <- unique(long_Data$Trait)
for (i in seq_along(traits)) {
  trait <- traits[i]
  trait_data <- subset(long_Data, Trait == trait)
  lowest_y_value <- min(trait_data$TraitValue, na.rm = TRUE)
  y_position <- lowest_y_value * 0.9
  
  plot <- ggplot(subset(long_Data, Trait == trait), aes(x = Species, y = TraitValue)) +
    geom_rain(alpha = .5, adjust = 0.3, point.args = list(alpha = 0),
              boxplot.args = list(alpha = 0, color = "white", outlier.shape = NA),
              violin.args = list(aes(fill = Species), size = 0.1, alpha = 0.66, adjust = 0.77),
              violin.args.pos = list(
                side = "r",
                width = 0.7, position = position_nudge(x = 0.22))) +
    geom_line(aes(group = Sample.source, color = Sample.source), position = position_dodge(width = 0.35), alpha = 0.5) +
    geom_point(aes(color = Sample.source), position = position_dodge(width = 0.35), size = 0.8, alpha = 0.5) +
    geom_text(aes(label = cld, y = y_position), size = 3) + # Add CLD labels 
    scale_fill_manual(values = SpeciesColorsPul) +
    scale_color_manual(values = all_strain_colors) +
    scale_x_discrete(labels = species_counts_veg$Species_label) +
    guides(color = FALSE) +
    theme_minimal() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 9)) +
    labs(y = trait)
    # Add an uppercase letter annotation to each plot
  plot <- plot + annotate("text", x = -Inf, y = Inf, label = LETTERS[i], hjust = -0.5, vjust = 1.5, size = 6, fontface = "bold")
  
  grob_list[[i]] <- plot
  #grob_list[[trait]] <- plot
}

# Arrange the plots side by side using grid.arrange
grid.arrange(grobs = grob_list, ncol = 3)

# Arrange the plots side by side using grid.arrange and save as PNG
png(filename = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/finalplots/plot_facet_rain_repr_2Pul_CLD.png", 
    width = 10, height = 12, units = "in", res = 300, bg = "white")
grid.arrange(grobs = grob_list, ncol = 3)
dev.off()
```



## Individual trait significance tests for results claims

### Levene's test for vegetative and reproductive traits
```{r}
# vegetative traits
# Reshape data to long format
vegetative_long <- vegetative_plots_data %>%
  pivot_longer(cols = all_of(includeVegTraits), names_to = "Trait", values_to = "TraitValue") %>%
  #pivot_longer(cols = includeTraitsWOcolour, names_to = "Trait", values_to = "TraitValue") %>%
  filter(!(Species.Code %in% c("Control", NA))) 
  #filter(!(Species %in% c("Control", "Aus", "Ppo", NA))) 

# Get unique traits
unique_traits <- unique(vegetative_long$Trait)

# Initialize a list to store results
veg_levene_test_results <- list()

# Loop through each trait and perform Levene's test
for (trait in unique_traits) {
  # Subset the dataframe for the current trait
  trait_data <- subset(vegetative_long, Trait == trait)
  
  # Perform Levene's test
  levene_test <- leveneTest(TraitValue ~ Species, data = trait_data)
  
  # Store the result in the list
  veg_levene_test_results[[trait]] <- levene_test
}

# Print results
veg_levene_test_results

# reproductive traits
# Reshape data to long format
reproductive_long <- reproductive_plots_data %>%
  pivot_longer(cols = all_of(includeReprTraits), names_to = "Trait", values_to = "TraitValue") %>%
  #pivot_longer(cols = includeTraitsWOcolour, names_to = "Trait", values_to = "TraitValue") %>%
  filter(!(Species %in% c("Control", "Aus", "Ppo", NA))) 

# Get unique traits
unique_traits <- unique(reproductive_long$Trait)

# Initialize a list to store results
repr_levene_test_results <- list()

# Loop through each trait and perform Levene's test
for (trait in unique_traits) {
  # Subset the dataframe for the current trait
  trait_data <- subset(reproductive_long, Trait == trait)
  
  # Perform Levene's test
  levene_test <- leveneTest(TraitValue ~ Species, data = trait_data)
  
  # Store the result in the list
  repr_levene_test_results[[trait]] <- levene_test
}

# Print results
repr_levene_test_results
```

### ANOVA on species means
```{r}
# Vegetative traits
# Get unique traits
unique_traits <- unique(vegetative_long$Trait)

# Initialize a list to store ANOVA results
anova_results <- list()

# Loop through each trait and perform ANOVA
for (trait in unique_traits) {
  # Subset the dataframe for the current trait
  trait_data <- vegetative_long %>%
    filter(Trait == trait)
  
  # Perform ANOVA
  anova_result <- aov(TraitValue ~ Species, data = trait_data)
  
  # Store the result in the list
  anova_results[[trait]] <- summary(anova_result)
}

# Print results
anova_results

# Reproductive traits
# Get unique traits
unique_traits <- unique(reproductive_long$Trait)

# Initialize a list to store ANOVA results
repr_anova_results <- list()

# Loop through each trait and perform ANOVA
for (trait in unique_traits) {
  # Subset the dataframe for the current trait
  trait_data <- reproductive_long %>%
    filter(Trait == trait)
  
  # Perform ANOVA
  anova_result <- aov(TraitValue ~ Species, data = trait_data)
  
  # Store the result in the list
  repr_anova_results[[trait]] <- summary(anova_result)
}

# Print results
repr_anova_results
```

### Tukey pairwise test
```{r}
#Vegetative traits
unique_traits <- unique(vegetative_long$Trait)

# Initialize a list to store Tukey's HSD results
vegetative_tukey_results <- list()

# Loop through each trait and perform Tukey's HSD test
for (trait in unique_traits) {
  # Subset the dataframe for the current trait
  trait_data <- vegetative_long %>%
    filter(Trait == trait)
  
  # Perform ANOVA
  anova_result <- aov(TraitValue ~ Species, data = trait_data)
  
  # Perform Tukey's HSD test
  tukey_result <- TukeyHSD(anova_result)
  
  # Store the result in the list
  vegetative_tukey_results[[trait]] <- tukey_result
}

# Print results for each trait
vegetative_tukey_results

#reproductive traits
unique_traits <- unique(reproductive_long$Trait)

# Initialize a list to store Tukey's HSD results
reproductive_tukey_results <- list()

# Loop through each trait and perform Tukey's HSD test
for (trait in unique_traits) {
  # Subset the dataframe for the current trait
  trait_data <- reproductive_long %>%
    filter(Trait == trait)
  
  # Perform ANOVA
  anova_result <- aov(TraitValue ~ Species, data = trait_data)
  
  # Perform Tukey's HSD test
  tukey_result <- TukeyHSD(anova_result)
  
  # Store the result in the list
  reproductive_tukey_results[[trait]] <- tukey_result
}

# Print results for each trait
reproductive_tukey_results
```

### Correlation water content and hardness
```{r}
# Checking normality (Shapiro-Wilk test)
shapiro.test(filtered_Data$Flush1_Hardness)
shapiro.test(filtered_Data$Flush1Fruitbody_water_content)

# Pearson correlation test
cor.test(filtered_Data$Flush1_Hardness, filtered_Data$Flush1Fruitbody_water_content, method = "kendall")

# Scatter plot
plot(filtered_Data$Flush1_Hardness, filtered_Data$Flush1Fruitbody_water_content, main = "Scatter plot of Hardness vs Water Content",
     xlab = "Hardness", ylab = "Water Content", pch = 19)
abline(lm(Flush1Fruitbody_water_content ~ Flush1_Hardness, data = filtered_Data), col = "blue") # Add a linear regression line
```

```{r}
# Assuming your data frame is named 'filtered_Data' and has a column 'Species'
# Split data by species
species_list <- c("Dja", "Pars", "Pul_NZ", "Pul_intl")

# Initialize an empty data frame to store correlation results
cor_results <- data.frame(Species = character(),
                          Method = character(),
                          Correlation = numeric(),
                          P_value = numeric(),
                          stringsAsFactors = FALSE)

# Perform correlation test for each species and store the results
for (species in species_list) {
  species_data <- filtered_Data %>% filter(Species == paste0(species))
  
  # Checking normality
  shapiro_hardness <- shapiro.test(species_data$Flush1_Hardness)
  shapiro_water <- shapiro.test(species_data$Flush1Fruitbody_water_content)
  
  # Determine which correlation test to use based on normality
  if (shapiro_hardness$p.value > 0.05 & shapiro_water$p.value > 0.05) {
    # Use Pearson correlation
    cor_test <- cor.test(species_data$Flush1_Hardness, species_data$Flush1Fruitbody_water_content, method = "pearson")
    method_used <- "Pearson"
  } else {
    # Use Spearman correlation
    cor_test <- cor.test(species_data$Flush1_Hardness, species_data$Flush1Fruitbody_water_content, method = "spearman")
    method_used <- "Spearman"
  }
  
  # Store the results in the data frame
  cor_results <- rbind(cor_results, data.frame(Species = species,
                                               Method = method_used,
                                               Correlation = cor_test$estimate,
                                               P_value = cor_test$p.value))
}

# Print correlation results
print(cor_results)

# Plotting the data
ggplot(filtered_Data, aes(x = Flush1_Hardness, y = Flush1Fruitbody_water_content, color = Species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Scatter plot of Hardness vs Water Content by Species",
       x = "Hardness",
       y = "Water Content") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

# Hypothesis 1: vegetative vs reproductive


### Variance components and confidence intervals of each trait
```{r}
# Initialize data frames to store results
results_CI_veg <- data.frame(Trait = character(), Within_species_variance_CI = character(), stringsAsFactors = FALSE)
results_CI_repr <- data.frame(Trait = character(), Within_species_variance_CI = character(), stringsAsFactors = FALSE)

# Function to calculate confidence intervals
calculate_ci <- function(variance, n, A, B) {
  lower_bound <- ((n - 1) * variance) / B
  upper_bound <- ((n - 1) * variance) / A
  return(c(lower_bound, upper_bound))
}

# Loop over each trait for vegetative dataset
for(trait in includeVegTraits) {
  formula <- as.formula(paste(trait, "~ (1 | Species/Sample.source) + (1 | shelf * level)"))
  
  model <- glmmTMB(formula, data = data_for_vegetative, family = gaussian())
  
  # Extract the variance components
  var_components <- VarCorr(model)
  
  # Within-species variance (Sample.source:Species)
  var_within_species <- var_components$cond$`Sample.source:Species`["(Intercept)", "(Intercept)"]
  
  # Extract the residual variance
  residual_variance <- sigma(model)^2
  
  # Extract other variance components and sum them up
  other_variances <- sapply(var_components$cond, function(x) x[1, 1])
  var_total <- sum(other_variances) + residual_variance
  
  # Calculate relative within-species variance
  relative_within_species_variance <- var_within_species / var_total
  
  # Calculate degrees of freedom
  df <- length(unique(data_for_vegetative$Sample.source))
  
  # Confidence level
  alpha <- 0.05
  
  # Chi-squared critical values
  A <- qchisq(alpha / 2, df = df - 1)
  B <- qchisq(1 - alpha / 2, df = df - 1)
  
  # Calculate confidence intervals for the relative variance component
  ci <- calculate_ci(relative_within_species_variance, df, A, B)
  ci_lower <- ci[1]
  ci_upper <- ci[2]
  
  # Create a result string
  ci_string <- paste0(round(relative_within_species_variance, 4), " (", round(ci_lower, 4), " - ", round(ci_upper, 4), ")")
  
  # Add to results data frame
  results_CI_veg <- rbind(results_CI_veg, data.frame(Trait = trait, Within_species_variance_CI = ci_string, stringsAsFactors = FALSE))
}

# Loop over each trait for reproductive dataset
for(trait in includeReprTraits) {
  formula <- as.formula(paste(trait, "~ (1 | Species/Sample.source) + (1 | shelf * level)"))
  
  model <- glmmTMB(formula, data = data_for_reproductive, family = gaussian())
  
  # Extract the variance components
  var_components <- VarCorr(model)
  
  # Within-species variance (Sample.source:Species)
  var_within_species <- var_components$cond$`Sample.source:Species`["(Intercept)", "(Intercept)"]
  
  # Extract the residual variance
  residual_variance <- sigma(model)^2
  
  # Extract other variance components and sum them up
  other_variances <- sapply(var_components$cond, function(x) x[1, 1])
  var_total <- sum(other_variances) + residual_variance
  
  # Calculate relative within-species variance
  relative_within_species_variance <- var_within_species / var_total
  
  # Calculate degrees of freedom
  df <- length(unique(data_for_reproductive$Sample.source))
  
  # Confidence level
  alpha <- 0.05
  
  # Chi-squared critical values
  A <- qchisq(alpha / 2, df = df - 1)
  B <- qchisq(1 - alpha / 2, df = df - 1)
  
  # Calculate confidence intervals for the relative variance component
  ci <- calculate_ci(relative_within_species_variance, df, A, B)
  ci_lower <- ci[1]
  ci_upper <- ci[2]
  
  # Create a result string
  ci_string <- paste0(round(relative_within_species_variance, 4), " (", round(ci_lower, 4), " - ", round(ci_upper, 4), ")")
  
  # Add to results data frame
  results_CI_repr <- rbind(results_CI_repr, data.frame(Trait = trait, Within_species_variance_CI = ci_string, stringsAsFactors = FALSE))
}

# Print the results
print(results_CI_veg)
print(results_CI_repr)
```


```{r}

# Combine veg and repr results
combined_results <- bind_rows(results_CI_veg, results_CI_repr)

# Separate the variance and confidence intervals into separate columns
combined_results <- combined_results %>%
  mutate(Within_species_variance = as.numeric(gsub(" .*", "", Within_species_variance_CI)),
         CI_lower = as.numeric(gsub(".*\\((.*) -.*", "\\1", Within_species_variance_CI)),
         CI_upper = as.numeric(gsub(".*- (.*)\\)", "\\1", Within_species_variance_CI)))

# Create the scatterplot
ggplot(combined_results, aes(x = Trait, y = Within_species_variance)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.2) +
  labs(title = "Within-species Variance with Confidence Intervals",
       x = "Trait",
       y = "Within-species Variance") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


## PCA - reproductive traits
```{r}
# Dataframe with reproductive observations, no categorical variables, removing rows with NAs
#data_for_reproductive_numeric <- na.omit(data_for_reproductive[, c(11:16, 19:26)])
data_for_reproductive_numeric <- na.omit(data_for_reproductive[, includeReprTraits])
#data_for_reproductive_numeric <- na.omit(data_for_reproductive)

standardized.reproductive <- decostand(data_for_reproductive_numeric[, -c(6)], method = "standardize")
pca_reproductive <- prcomp(standardized.reproductive, scale=T)
print(biplot(pca_reproductive))
summary(pca_reproductive)

# Extract the loadings
loadings_matrix <- pca_reproductive$rotation

# Calculate the magnitude of loadings for each variable 
loadings_magnitude <- sqrt(rowSums(loadings_matrix^2))

# Identify the 6 most prominent variables
top_6_vars <- names(sort(loadings_magnitude, decreasing = TRUE)[1:6])

# Subset the PCA object to keep only the top 6 variables
pca_reproductive$rotation <- loadings_matrix[top_6_vars, ]

# Now, plot the biplot
biplot(pca_reproductive)
```

## PCA - vegetative traits
```{r}
# Dataframe with reproductive observations, no categorical variables, removing rows with NAs
#data_for_vegetative_numeric <- na.omit(data_for_vegetative[, c(10, 17, 18)])
data_for_vegetative_numeric <- na.omit(data_for_vegetative[, includeVegTraits])

standardized.vegetative <- decostand(data_for_vegetative_numeric, method = "standardize")
pca_vegetative <- prcomp(standardized.vegetative, scale=T)
print(biplot(pca_vegetative))
summary(pca_vegetative)
```

## Variance components using glmmTMB
### Function for extracting variance from glmmTMB models
```{r}
glmm_var <- function(model, number_of_terms) { #specify number of random effects terms (usually 3 for my data)
  model_variance <- glmmTMB::VarCorr(model) #extract variance components from model
    df <- data.frame(name = rep(NA, number_of_terms + 1),
                   variance = rep(NA, number_of_terms + 1)) # create empty dataframe with i rows
  for (i in 1:number_of_terms) {
    category <- names(model_variance$cond)[i] # extract name of category of variance
    #print(paste("Category:", category))  # Added for debugging
    df[i, 1] <- category
    df[(i+1), 1] <- "Residual"
    value <- model_variance$cond[[category]][1] # extract value of variance for i-th category
    #print(paste("Value:", value))  # Added for debugging
    df[i, 2] <- value
    df[(i+1), 2] <- attr(model_variance$cond, "sc")^2 # Residual Variance
    #print(paste("Residual Variance for row", (i+1), ":", df[(i+1), 2])) # Echo back the value

  }
  return(df)
}

# Function to calculate absolute and relative variances
glmm_varcomp <- function(model, number_of_terms) { 
  # Extract variance components from model
  model_variance <- glmmTMB::VarCorr(model) 

  # Create empty dataframe with appropriate rows
  df <- data.frame(name = rep(NA, number_of_terms + 1),
                   variance = rep(NA, number_of_terms + 1),
                   varcomp = rep(NA, number_of_terms + 1)) 

  # Calculate absolute variances
  for (i in 1:number_of_terms) {
    category <- names(model_variance$cond)[i] 
    df[i, 1] <- category 
    df[(i + 1), 1] <- "Residual"
    value <- model_variance$cond[[category]][1]  
    df[i, 2] <- value 
    df[(i + 1), 2] <- attr(model_variance$cond, "sc")^2 
  }

  # Calculate relative variances
  df$varcomp <- df$variance / sum(df$variance)

  return(df)
}

  
# Function to calculate relative variances
relative_variance <- function(data) {
  model_variance <- data
# Loop through each column in the dataframe
for (col_name in names(model_variance)) {
  # Check if the column is numeric
  if (is.numeric(model_variance[[col_name]])) {
    # Normalize the column by dividing each value by the sum of the column, handling NA values
    model_variance[[col_name]] <- model_variance[[col_name]] / sum(model_variance[[col_name]], na.rm = TRUE)
  }

}
  # View the updated dataframe
return(model_variance)

}
```


### PCA variance components
```{r}
## One Pul species
vegetative_pcoa1_glmmTMB <- glmmTMB(pca_vegetative$x[,1] ~ (1 | Species.Code/Sample.source) + (1 | shelfxlevel), data = data_for_vegetative, family = gaussian())
vegetative_pcoa1_varcomp <- format(glmm_varcomp(vegetative_pcoa1_glmmTMB,3), scientific = F) #extract variance components, absolute and relative
vegetative_pcoa2_glmmTMB <- glmmTMB(pca_vegetative$x[,2] ~ (1 | Species.Code/Sample.source) + (1 | shelfxlevel), data = data_for_vegetative, family = gaussian())
vegetative_pcoa2_varcomp <- format(glmm_varcomp(vegetative_pcoa2_glmmTMB,3), scientific = F)

reproductive_pcoa1_glmmTMB <- glmmTMB(pca_reproductive$x[,1] ~ (1 | Species.Code/Sample.source) + (1 | shelfxlevel), data = data_for_reproductive, family = gaussian())
reproductive_pcoa1_varcomp <- format(glmm_varcomp(reproductive_pcoa1_glmmTMB,3), scientific = F)
reproductive_pcoa2_glmmTMB <- glmmTMB(pca_reproductive$x[,2] ~ (1 | Species.Code/Sample.source) + (1 | shelfxlevel), data = data_for_reproductive, family = gaussian())
reproductive_pcoa2_varcomp <- format(glmm_varcomp(reproductive_pcoa2_glmmTMB,3), scientific = F)

### Varcomp summary (relative variances)
PCoA_varcomp_summary <- vegetative_pcoa1_varcomp[, c(1,3)]
names(PCoA_varcomp_summary)[names(PCoA_varcomp_summary) == "varcomp"] <- "vegetative_pcoa1"
new_levels <- c("Between strains", "Between species", "Shelf effect", "Within strains")
PCoA_varcomp_summary$name <- factor(new_levels, levels = new_levels)
column_name <- "vegetative_pcoa2"
PCoA_varcomp_summary[[column_name]] <- vegetative_pcoa2_varcomp[[3]] # Append the column to the new dataframe
column_name <- "reproductive_pcoa1"
PCoA_varcomp_summary[[column_name]] <- reproductive_pcoa1_varcomp[[3]] # Append the column to the new dataframe
column_name <- "reproductive_pcoa2"
PCoA_varcomp_summary[[column_name]] <- reproductive_pcoa2_varcomp[[3]] # Append the column to the new dataframe

#####################################
## Two Pul species
Pul_vegetative_pcoa1_glmmTMB <- glmmTMB(pca_vegetative$x[,1] ~ (1 | Species/Sample.source) + (1 | shelf * level), data = data_for_vegetative, family = gaussian())
Pul_vegetative_pcoa1_varcomp <- format(glmm_varcomp(Pul_vegetative_pcoa1_glmmTMB,5), scientific = F) #extract variance components, absolute and relative
Pul_vegetative_pcoa2_glmmTMB <- glmmTMB(pca_vegetative$x[,2] ~ (1 | Species/Sample.source) + (1 | shelf * level), data = data_for_vegetative, family = gaussian())
Pul_vegetative_pcoa2_varcomp <- format(glmm_varcomp(Pul_vegetative_pcoa2_glmmTMB,5), scientific = F)

Pul_reproductive_pcoa1_glmmTMB <- glmmTMB(pca_reproductive$x[,1] ~ (1 | Species/Sample.source) + (1 | shelf * level), data = data_for_reproductive, family = gaussian())
Pul_reproductive_pcoa1_varcomp <- format(glmm_varcomp(Pul_reproductive_pcoa1_glmmTMB,5), scientific = F)
Pul_reproductive_pcoa2_glmmTMB <- glmmTMB(pca_reproductive$x[,2] ~ (1 | Species/Sample.source) + (1 | shelf * level), data = data_for_reproductive, family = gaussian())
Pul_reproductive_pcoa2_varcomp <- format(glmm_varcomp(Pul_reproductive_pcoa2_glmmTMB,5), scientific = F)

### Varcomp summary (relative variances)
Pul_PCoA_varcomp_summary <- Pul_vegetative_pcoa1_varcomp[, c(1,3)]
names(Pul_PCoA_varcomp_summary)[names(Pul_PCoA_varcomp_summary) == "varcomp"] <- "vegetative_pc1"
#new_levels <- c("Between strains", "Between species", "Shelf effect", "Within strains")
#new_levels <- c("Between strains", "Between species", "Shelf effect", "Level effect", "Shelf * level", "Water content", "Within strains")
new_levels <- c("Between strains", "Between species", "Shelf effect", "Level effect", "Shelf  level", "Within strains")
Pul_PCoA_varcomp_summary$name <- factor(new_levels, levels = new_levels)
column_name <- "vegetative_pc2"
Pul_PCoA_varcomp_summary[[column_name]] <- Pul_vegetative_pcoa2_varcomp[[3]] # Append the column to the new dataframe
column_name <- "reproductive_pc1"
Pul_PCoA_varcomp_summary[[column_name]] <- Pul_reproductive_pcoa1_varcomp[[3]] # Append the column to the new dataframe
column_name <- "reproductive_pc2"
Pul_PCoA_varcomp_summary[[column_name]] <- Pul_reproductive_pcoa2_varcomp[[3]] # Append the column to the new dataframe

# Manually sort
new_sorting <- rev(c("Between species", "Between strains", "Within strains", "Shelf  level", "Shelf effect", "Level effect"))
Pul_PCoA_varcomp_summary$name <- factor(Pul_PCoA_varcomp_summary$name, levels = new_sorting)

```


#### Stacked barplots of variance components
```{r}
# color palette for chart
colors_barplot <- c("#009E73", "#253998", "#FFEA97", "#44C3FF", "#a9d574", "#fa753c")

## Two species
PCA_varcomp_summary_long <- Pul_PCoA_varcomp_summary %>%
  pivot_longer(cols = 2:5, names_to = "PCA", values_to = "value") %>%
  mutate(value = as.numeric(value)) # Ensure 'value' is numeric

colors_named <- setNames(colors_barplot, unique(PCA_varcomp_summary_long$name))

# Now, create the stacked barplot
plot_hyp1varcomp <- ggplot(PCA_varcomp_summary_long, aes(x = PCA, y = value, fill = name)) + 
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  scale_fill_manual(values = colors_named) +
  labs(x = "PCA Analysis", 
       y = "Relative variance", 
       fill = "Variance component") +
  theme(text = element_text(size = 10), axis.text.x = element_text(angle = 45, hjust = 0.8), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.title.x = element_blank())

print(plot_hyp1varcomp)

ggsave(filename = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/finalplots/plot.hyp1varcomp.png", plot = plot_hyp1varcomp, width = 3, height = 2, dpi = 300, bg = "white", scale = 2)
```

### Log likelihood ratio test
```{r}
## Comparing the full model against a reduced model, in which either Species or Strain is removed as a random effect 
# Full model
full_model <- glmmTMB(pca_reproductive$x[,1] ~ (1 | Species/Sample.source) + (1 | shelfxlevel), 
                      data = data_for_reproductive, family = gaussian())

# Reduced model (example: removing between-species variance)
reduced_model <- glmmTMB(pca_reproductive$x[,1] ~ (1 | Species) + (1 | shelfxlevel), 
                         data = data_for_reproductive, family = gaussian())

# Likelihood ratio test
print(anova(full_model, reduced_model))

```

#### Model comparison function to create table
```{r}
### Currently uses two Pul clades

compare_models <- function(data, response_vars) {
  # Initialize an empty data frame to store results
  results <- data.frame(Response=character(),
                        Model=character(),
                        AIC_delta=numeric(),
                        Chisq=numeric(),
                        P=numeric(),
                        stringsAsFactors=FALSE)

  for (var in response_vars) {
    # Full model
    full_model_formula <- as.formula(paste0(var, " ~ (1 | Species/Sample.source) + (1 | shelfxlevel)"))
    full_model <- glmmTMB(full_model_formula, data = data, family = gaussian())

    # Reduced model 1: Sample.source only
    reduced_model_1_formula <- as.formula(paste0(var, " ~ (1 | Sample.source) + (1 | shelfxlevel)"))
    reduced_model_1 <- glmmTMB(reduced_model_1_formula, data = data, family = gaussian())

    # Reduced model 2: Species.Code only
    reduced_model_2_formula <- as.formula(paste0(var, " ~ (1 | Species) + (1 | shelfxlevel)"))
    reduced_model_2 <- glmmTMB(reduced_model_2_formula, data = data, family = gaussian())

    # Model comparison 1
    comparison_1 <- anova(full_model, reduced_model_1)

    # Model comparison 2
    comparison_2 <- anova(full_model, reduced_model_2)

    # Collect results directly, avoiding the use of `tidy()`
    results <- rbind(results, 
                     data.frame(Response = var,
                                Model = "Strain only",
                                AIC_delta = comparison_1$AIC[2] - comparison_1$AIC[1],
                                Chisq = comparison_1$`Chisq`[2],
                                P = comparison_1$`Pr(>Chisq)`[2]),
                     data.frame(Response = var,
                                Model = "Species only",
                                AIC_delta = comparison_2$AIC[2] - comparison_2$AIC[1],
                                Chisq = comparison_2$`Chisq`[2],
                                P = comparison_2$`Pr(>Chisq)`[2]))
  }
  
  return(results)
}

# Running function
response_vars <- c("pca_reproductive$x[,1]", "pca_reproductive$x[,2]")
results_table_reproductive <- compare_models(data_for_reproductive, response_vars)
response_vars <- c("pca_vegetative$x[,1]", "pca_vegetative$x[,2]")
results_table_vegetative <- compare_models(data_for_vegetative, response_vars)

# Renaming entries
results_table_vegetative$Model <- as.factor(results_table_vegetative$Model)
# Rename the levels of the Model factor
levels(results_table_vegetative$Model)[levels(results_table_vegetative$Model) == "Strain only"] <- "Species effect"
levels(results_table_vegetative$Model)[levels(results_table_vegetative$Model) == "Species only"] <- "Strain effect"

results_table_vegetative$Response <- as.factor(results_table_vegetative$Response)
# Rename the levels of the Response factor
levels(results_table_vegetative$Response)[levels(results_table_vegetative$Response) == "pca_vegetative$x[,1]"] <- "PC1"
levels(results_table_vegetative$Response)[levels(results_table_vegetative$Response) == "pca_vegetative$x[,2]"] <- "PC2"

results_table_reproductive$Model <- as.factor(results_table_reproductive$Model)
# Rename the levels of the Model factor
levels(results_table_reproductive$Model)[levels(results_table_reproductive$Model) == "Strain only"] <- "Species effect"
levels(results_table_reproductive$Model)[levels(results_table_reproductive$Model) == "Species only"] <- "Strain effect"

results_table_reproductive$Response <- as.factor(results_table_reproductive$Response)
# Rename the levels of the Response factor
levels(results_table_reproductive$Response)[levels(results_table_reproductive$Response) == "pca_reproductive$x[,1]"] <- "PC1"
levels(results_table_reproductive$Response)[levels(results_table_reproductive$Response) == "pca_reproductive$x[,2]"] <- "PC2"

# Create table for publication
ft_veg <- flextable(results_table_vegetative)
ft_veg <- set_formatter(ft_veg, P = function(x) {
    formatC(x, format = "e", digits = 2)
  })
ft_veg <- set_formatter(ft_veg, Chisq = function(x) {
    formatC(x, format = "f", digits = 2)
  })
ft_veg <- set_formatter(ft_veg, AIC_delta = function(x) {
    formatC(x, format = "f", digits = 2)
  })
ft_veg <- autofit(ft_veg)

ft_repr <- flextable(results_table_reproductive)
ft_repr <- set_formatter(ft_repr, P = function(x) {
    formatC(x, format = "e", digits = 2)
  })
ft_repr <- set_formatter(ft_repr, Chisq = function(x) {
    formatC(x, format = "f", digits = 2)
  })
ft_repr <- set_formatter(ft_repr, AIC_delta = function(x) {
    formatC(x, format = "f", digits = 2)
  })
ft_repr <- autofit(ft_repr)


save_as_docx(
  "Vegetative traits" = ft_veg, "Reproductive traits" = ft_repr, 
  path = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/flextable_modelcomp.docx")

```


## Betadisp hyp 1 testing
### Function to calculate centroids manually
```{r}
    centroid_calc <- function(vec, group) { # vec = distance matrix
        cent <- apply(vec, 2, function(x, group) tapply(x, INDEX = group, 
            FUN = mean), group = group)
        if (!is.matrix(cent)) {
            cent <- matrix(cent, nrow = 1, dimnames = list(as.character(levels(group)), 
                paste0("Dim", seq_len(NCOL(vec)))))
        }
        cent
    }

#centroid_strains <- centroid_calc(vegdist(reproductive_results, method = "gower"), data_for_reproductive$Sample.source)


## Modified betadisper to calculate centroids and group distances for a second  group - this seems a bit overkill as I could have just created two models, one per group, and plotted them on top of each other. I tested it both ways and results are the same.
betadisper2 <- function (d, group, group2, type = c("median", "centroid"), bias.adjust = FALSE, 
    sqrt.dist = FALSE, add = FALSE) 
{
    dblcen <- function(x, na.rm = TRUE) {
        cnt <- colMeans(x, na.rm = na.rm)
        x <- sweep(x, 2L, cnt, check.margin = FALSE)
        cnt <- rowMeans(x, na.rm = na.rm)
        sweep(x, 1L, cnt, check.margin = FALSE)
    }
    spatialMed <- function(vectors, group, pos) {
        axes <- seq_len(NCOL(vectors))
        spMedPos <- vegan:::ordimedian(vectors, group, choices = axes[pos])
        spMedNeg <- vegan:::ordimedian(vectors, group, choices = axes[!pos])
        cbind(spMedPos, spMedNeg)
    }
    centroidFUN <- function(vec, group) {
        cent <- apply(vec, 2, function(x, group) tapply(x, INDEX = group, 
            FUN = mean), group = group)
        if (!is.matrix(cent)) {
            cent <- matrix(cent, nrow = 1, dimnames = list(as.character(levels(group)), 
                paste0("Dim", seq_len(NCOL(vec)))))
        }
        cent
    }
    Resids <- function(x, c) {
        if (is.matrix(c)) 
            d <- x - c
        else d <- sweep(x, 2, c)
        rowSums(d^2)
    }
    TOL <- sqrt(.Machine$double.eps)
    if (!inherits(d, "dist")) 
        stop("distances 'd' must be a 'dist' object")
    if (any(d < -TOL, na.rm = TRUE)) 
        stop("dissimilarities 'd' must be non-negative")
    if (sqrt.dist) 
        d <- sqrt(d)
    if (is.logical(add) && isTRUE(add)) 
        add <- "lingoes"
    if (is.character(add)) {
        add <- match.arg(add, c("lingoes", "cailliez"))
        if (add == "lingoes") {
            ac <- addLingoes(as.matrix(d))
            d <- sqrt(d^2 + 2 * ac)
        }
        else if (add == "cailliez") {
            ac <- addCailliez(as.matrix(d))
            d <- d + ac
        }
    }
    if (missing(type)) 
        type <- "median"
    type <- match.arg(type)
    group <- if (!is.factor(group)) {
        as.factor(group)
    }
    else {
        droplevels(group, exclude = NA)
    }
    group2 <- if (!is.factor(group2)) {
        as.factor(group2)
    } else {
        droplevels(group2, exclude = NA)
    }
    n <- attr(d, "Size")
    x <- matrix(0, ncol = n, nrow = n)
    x[row(x) > col(x)] <- d^2
    labs <- attr(d, "Labels")
    if (any(gr.na <- is.na(group))) {
        group <- group[!gr.na]
        x <- x[!gr.na, !gr.na]
        n <- n - sum(gr.na)
        labs <- labs[!gr.na]
        message("missing observations due to 'group' removed")
    }
    if (any(x.na <- apply(x, 1, function(x) any(is.na(x))))) {
        x <- x[!x.na, !x.na]
        group <- group[!x.na]
        n <- n - sum(x.na)
        labs <- labs[!x.na]
        message("missing observations due to 'd' removed")
    }
    x <- x + t(x)
    x <- dblcen(x)
    e <- eigen(-x/2, symmetric = TRUE)
    vectors <- e$vectors
    eig <- e$values
    eig <- eig[(want <- abs(eig) > max(TOL, TOL * eig[1L]))]
    vectors <- vectors[, want, drop = FALSE] %*% diag(sqrt(abs(eig)), 
        nrow = length(eig))
    pos <- eig > 0
    centroids <- switch(type, centroid = centroidFUN(vectors, 
        group), median = spatialMed(vectors, group, pos))
    centroids2 <- switch(type, 
                         centroid = centroidFUN(vectors, group2), 
                         median = spatialMed(vectors, group2, pos))
    dist.pos <- Resids(vectors[, pos, drop = FALSE], centroids[group, 
        pos, drop = FALSE])
    dist.neg <- 0
    if (any(!pos)) 
        dist.neg <- Resids(vectors[, !pos, drop = FALSE], centroids[group, 
            !pos, drop = FALSE])
    if (any(dist.neg > dist.pos)) {
        warning("some squared distances are negative and changed to zero")
        zij <- Re(sqrt(as.complex(dist.pos - dist.neg)))
    }
    else {
        zij <- sqrt(dist.pos - dist.neg)
    }
    if (bias.adjust) {
        n.group <- as.vector(table(group))
        zij <- zij * sqrt(n.group[group]/(n.group[group] - 1))
    }
    grp.zij <- tapply(zij, group, "mean")
    # Calculate distances for group2
    dist.pos2 <- Resids(vectors[, pos, drop = FALSE], centroids2[group2, pos, drop = FALSE])
    dist.neg2 <- 0
    if (any(!pos)) 
        dist.neg2 <- Resids(vectors[, !pos, drop = FALSE], centroids2[group2, !pos, drop = FALSE])
    if (any(dist.neg2 > dist.pos2)) {
        warning("some squared distances for group2 are negative and changed to zero")
        zij2 <- Re(sqrt(as.complex(dist.pos2 - dist.neg2)))
    }
    else {
        zij2 <- sqrt(dist.pos2 - dist.neg2)
    }
    if (bias.adjust) {
        n.group2 <- as.vector(table(group2))
        zij2 <- zij2 * sqrt(n.group2[group2]/(n.group2[group2] - 1))
    }

    # Compute mean distances for each level in group2
    grp.zij2 <- tapply(zij2, group2, "mean")
    if (any(want)) 
        colnames(vectors) <- names(eig) <- paste("PCoA", seq_along(eig), 
            sep = "")
    if (is.matrix(centroids)) 
        colnames(centroids) <- names(eig)
    else names(centroids) <- names(eig)
    rownames(vectors) <- names(zij) <- labs
    retval <- list(eig = eig, vectors = vectors, distances = zij, 
        group = group, centroids = centroids, group2 = group2, centroids2 = centroids2, # Include group2 and its centroids
        group.distances = grp.zij, group2.distances = grp.zij2, call = match.call())
    class(retval) <- "betadisper"
    attr(retval, "method") <- attr(d, "method")
    attr(retval, "type") <- type
    attr(retval, "bias.adjust") <- bias.adjust
    retval
}

centroid2_plot <- function (x, axes = c(1, 2), cex = 0.7, pch = seq_len(ng), col = NULL, 
    lty = "solid", lwd = 1, hull = TRUE, ellipse = FALSE, conf, 
    segments = TRUE, seg.col = "grey", seg.lty = lty, seg.lwd = lwd, 
    label = TRUE, label.cex = 1, ylab, xlab, main, sub, ...) 
{
    localAxis <- function(..., col, bg, pch, cex, lty, lwd) axis(...)
    localBox <- function(..., col, bg, pch, cex, lty, lwd) box(...)
    localTitle <- function(..., col, bg, pch, cex, lty, lwd) title(...)
    Ellipse <- function(scrs, centres, conf, col, lty, lwd, ...) {
        mat <- cov.wt(scrs, center = centres)
        if (mat$n.obs == 1) 
            mat$cov[] <- 0
        xy <- if (mat$n.obs > 1) {
            veganCovEllipse(mat$cov, mat$center, conf)
        }
        else {
            scrs
        }
        ordiArgAbsorber(xy, FUN = lines, col = col, lty = lty, 
            lwd = lwd, ...)
    }
    if (missing(main)) 
        main <- deparse(substitute(x))
    if (missing(sub)) 
        sub <- paste("method = \"", attr(x, "method"), "\"", 
            sep = "")
    if (missing(xlab)) 
        xlab <- paste("PCoA", axes[1])
    if (missing(ylab)) 
        ylab <- paste("PCoA", axes[2])
    t <- if (missing(conf)) {
        1
    }
    else {
        sqrt(qchisq(conf, df = 2))
    }
    g <- scores(x, choices = axes)
    ng <- length(levels(x$group2))
    lev <- levels(x$group2)
    if (is.null(col)) {
        col <- palette()
    }
    col <- rep_len(col, ng)
    seg.col <- rep_len(seg.col, ng)
    plot(g$sites, asp = 1, type = "n", axes = FALSE, ann = FALSE, 
        ...)
    if (is.matrix(g$centroids2)) {
        for (i in seq_along(lev)) {
            curlev <- lev[i]
            take <- x$group2 == curlev
            j <- which(lev == curlev)
            if (segments) {
                segments(g$centroids2[j, 1L], g$centroids2[j, 2L], 
                  g$sites[take, 1L], g$sites[take, 2L], col = seg.col[i], 
                  lty = seg.lty, lwd = seg.lwd)
            }
            if (hull) {
                ch <- chull(g$sites[take, , drop = FALSE])
                ch <- c(ch, ch[1])
                lines(x$vectors[take, axes, drop = FALSE][ch, 
                  ], col = col[i], lty = lty, lwd = lwd, ...)
            }
            if (ellipse) {
                Ellipse(g$sites[take, , drop = FALSE], centres = g$centroids2[j, 
                  ], conf = t, col = col[i], lty = lty, lwd = lwd, 
                  ...)
            }
            points(g$centroids2[j, , drop = FALSE], pch = 16, 
                cex = 1, col = col[i], ...)
        }
    }
    else {
        if (segments) {
            segments(g$centroids[, 1L], g$centroids[, 2L], g$sites[, 
                1L], g$sites[, 2L], col = seg.col, lty = seg.lty, 
                ...)
        }
        if (hull) {
            ch <- chull(g$sites)
            ch <- c(ch, ch[1])
            lines(x$vectors[, axes][ch, ], col = col[1L], lty = lty, 
                lwd = lwd, ...)
        }
        if (ellipse) {
            Ellipse(g$sites, centres = g$centroids2, conf = t, 
                col = col[1L], lty = lty, lwd = lwd, ...)
        }
        points(g$centroids2[, 1L], g$centroids2[, 2L], pch = 16, 
            cex = 1, col = col[1L], ...)
    }
    points(g$sites, pch = pch[x$group2], cex = cex, col = col[x$group2], 
        ...)
    if (label) {
        ordilabel(x, display = "centroids2", choices = axes, cex = label.cex, 
            col = col)
    }
    localTitle(main = main, xlab = xlab, ylab = ylab, sub = sub, 
        ...)
    localAxis(1, ...)
    localAxis(2, ...)
    localBox(...)
    class(g) <- "ordiplot"
    invisible(g)
}
```

```{r}
# Plot.betadisper modified to draw ellipse polygons, not just outlines

plot_ellipse <- function (x, axes = c(1, 2), cex = 0.7, pch = seq_len(ng), col = NULL, 
    lty = "solid", lwd = 1, hull = TRUE, ellipse = FALSE, conf, 
    segments = TRUE, seg.col = "grey", seg.lty = lty, seg.lwd = lwd, 
    label = TRUE, label.cex = 1, ylab, xlab, main, sub, ...) 
{
    localAxis <- function(..., col, bg, pch, cex, lty, lwd) axis(...)
    localBox <- function(..., col, bg, pch, cex, lty, lwd) box(...)
    localTitle <- function(..., col, bg, pch, cex, lty, lwd) title(...)
    Ellipse <- function(scrs, centres, conf, col, lty, lwd, ...) {
        mat <- cov.wt(scrs, center = centres)
        if (mat$n.obs == 1) 
            mat$cov[] <- 0
        xy <- if (mat$n.obs > 1) {
            vegan:::veganCovEllipse(mat$cov, mat$center, conf)
        }
        else {
            scrs
        }
        transparent_col <- adjustcolor(col, alpha.f = 0.5)
        polygon(xy, col = transparent_col, border = col, lwd = lwd, ...)
        #ordiArgAbsorber(xy, FUN = lines, col = col, lty = lty, 
        #    lwd = lwd, ...)
    }
    if (missing(main)) 
        main <- deparse(substitute(x))
    if (missing(sub)) 
        sub <- paste("method = \"", attr(x, "method"), "\"", 
            sep = "")
    if (missing(xlab)) 
        xlab <- paste("PCoA", axes[1])
    if (missing(ylab)) 
        ylab <- paste("PCoA", axes[2])
    t <- if (missing(conf)) {
        1
    }
    else {
        sqrt(qchisq(conf, df = 2))
    }
    g <- scores(x, choices = axes)
    ng <- length(levels(x$group))
    lev <- levels(x$group)
    if (is.null(col)) {
        col <- palette()
    }
    col <- rep_len(col, ng)
    seg.col <- rep_len(seg.col, ng)
    plot(g$sites, asp = 1, type = "n", axes = FALSE, ann = FALSE, 
        ...)
    if (is.matrix(g$centroids)) {
        for (i in seq_along(lev)) {
            curlev <- lev[i]
            take <- x$group == curlev
            j <- which(lev == curlev)
            if (segments) {
                segments(g$centroids[j, 1L], g$centroids[j, 2L], 
                  g$sites[take, 1L], g$sites[take, 2L], col = seg.col[i], 
                  lty = seg.lty, lwd = seg.lwd)
            }
            if (hull) {
                ch <- chull(g$sites[take, , drop = FALSE])
                ch <- c(ch, ch[1])
                lines(x$vectors[take, axes, drop = FALSE][ch, 
                  ], col = col[i], lty = lty, lwd = lwd, ...)
            }
            if (ellipse) {
                Ellipse(g$sites[take, , drop = FALSE], centres = g$centroids[j, 
                  ], conf = t, col = col[i], lty = lty, lwd = lwd, 
                  ...)
            }
            #points(g$centroids[j, , drop = FALSE], pch = 16, 
            #    cex = 1, col = col[i], ...)
        }
    }
    else {
        if (segments) {
            segments(g$centroids[, 1L], g$centroids[, 2L], g$sites[, 
                1L], g$sites[, 2L], col = seg.col, lty = seg.lty, 
                ...)
        }
        if (hull) {
            ch <- chull(g$sites)
            ch <- c(ch, ch[1])
            lines(x$vectors[, axes][ch, ], col = col[1L], lty = lty, 
                lwd = lwd, ...)
        }
        if (ellipse) {
            Ellipse(g$sites, centres = g$centroids, conf = t, 
                col = col[1L], lty = lty, lwd = lwd, ...)
        }
        points(g$centroids[, 1L], g$centroids[, 2L], pch = 16, 
            cex = 1, col = col[1L], ...)
    }
    points(g$sites, pch = pch[x$group], cex = cex, col = col[x$group], 
        ...)
    if (label) {
        ordilabel(x, display = "centroids", choices = axes, cex = label.cex, 
            col = col)
    }
    localTitle(main = main, xlab = xlab, ylab = ylab, sub = sub, 
        ...)
    localAxis(1, ...)
    localAxis(2, ...)
    localBox(...)
    class(g) <- "ordiplot"
    invisible(g)
}
```


### Betadisp reproductive species level
```{r}
### Change dataframe to average per strain
# For reproductive_results
reproductive_results$ID <- rownames(reproductive_results)

# For data_for_reproductive (assuming Species.Code is here)
data_for_reproductive$ID <- rownames(data_for_reproductive)

# Merge to add Species.Code to reproductive_results based on row names (now in ID)
reproductive_results_with_species <- merge(reproductive_results, data_for_reproductive[, c("ID", "Sample.source", "Species")], by = "ID", all.x = TRUE)

# Remove the 'ID' column before aggregation if it's not needed for the analysis
reproductive_results_with_species$ID <- NULL
reproductive_results$ID <- NULL
data_for_reproductive$ID <- NULL

# Aggregate the means for each species
strain_means <- aggregate(. ~ Species/Sample.source, data = reproductive_results_with_species, FUN=mean)

## Reproductive traits
# Between strains
mod_disp <- betadisper(vegdist(reproductive_results, method = "euclidean"), data_for_reproductive$Species)
disp_test <- permutest(mod_disp, permutations=999)
print(disp_test)

# Calculate centroids of strains with my custom betadisper2 function creating centroids2
mod_disp2 <- betadisper2(vegdist(reproductive_results, method = "euclidean"), data_for_reproductive$Species, data_for_reproductive$Sample.source)

# To directly inspect the mean distances for each group
means <- aggregate(mod_disp$distances ~ mod_disp$group, FUN=mean)
print(means)
print(avg_strain_dispersion <- mean(means$`mod_disp$distances`))

# Pairwise comparison of centroids (variances) using Tukey HSD
tukey_test <- TukeyHSD(mod_disp)
print(tukey_test)

# Plotting
boxplot(mod_disp)

xlim_values <- c(-0.585, 0.32) #c(min(pcoa_vegetative_ape$vectors[,1]), max(pcoa_vegetative_ape$vectors[,1]))
ylim_values <- c(-0.20, 0.20)
species_to_pch <- setNames(c(0, 1, 2, 6, 5, 23, 3), levels(data_for_vegetative$Species)) 
species_to_pch_filled <- setNames(c(22, 21, 24, 25, 23, 5, 3), levels(data_for_vegetative$Species)) 

sample_source_to_species <- data_for_reproductive$Species[match(unique(data_for_reproductive$Sample.source), data_for_reproductive$Sample.source)]
#SpeciesColorsPul <- setNames(SpeciesColorsPul, levels(sample_source_to_species))
#SpeciesColorsPul <- factor(SpeciesColorsPul, levels(sample_source_to_species))

# Species plot
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.reproductive.species_hull.png", width = 600, height = 450)
plot(mod_disp, pch = NA, ellipse = F, hull = T, label = F, segments = FALSE, cex = 1, col = "lightgray")
#plot(mod_disp$centroids[,1], mod_disp$centroids[,2], xlim = xlim_values, ylim = ylim_values)
points(mod_disp2$vectors[,1], mod_disp2$vectors[,2], pch = species_to_pch[data_for_reproductive$Species], cex = 1.5, col = "lightgray") # replicates - raw datapoints
#plot(mod_disp, pch = NA, ellipse = F, hull = FALSE, label = F, segments = FALSE, cex = 2, col = "grey")
#points(mod_disp2$centroids2[,1], mod_disp2$centroids2[,2], pch = species_to_pch_filled[strain_means$Species], cex = 2, bg = "lightgray", col = NA) # strain centroids
#ordihull(mod_disp, groups = data_for_reproductive$Species, col = "grey", draw = "polygon")
points(mod_disp2$centroids[,1], mod_disp2$centroids[,2], pch = species_to_pch_filled, cex = 3, bg = SpeciesColorsPul, col = NA) #species centroids
```

#### Species plot
```{r}
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.reproductive.species_hull_noleg.png", width = 600, height = 450)

# Base plot
plot(mod_disp, pch = NA, ellipse = F, hull = T, label = F, segments = FALSE, cex = 1, col = "lightgray", main = "", xlab = "", ylab = "")
title(xlab = "PCoA 1", ylab = "PCoA 2", cex.lab = 1.5)

# Plot raw datapoints (replicates)
points(mod_disp2$vectors[,1], mod_disp2$vectors[,2], pch = species_to_pch[data_for_reproductive$Species], cex = 1.5, col = "lightgray")

# Plot species centroids
points(mod_disp2$centroids[,1], mod_disp2$centroids[,2], pch = species_to_pch_filled, cex = 3, bg = SpeciesColorsPul, col = NA)

dev.off()

specieslabels_reordered <- specieslabels[1:2]
specieslabels_reordered[3] <- specieslabels[4]
specieslabels_reordered[4] <- specieslabels[3]

# Create grob object for the legend

grob_legend <- legendGrob(labels = specieslabels_reordered,
                          pch = species_to_pch_filled[unique(data_for_reproductive$Species)],
                          gp = gpar(col = SpeciesColorsPul[unique(data_for_reproductive$Species)], fill = SpeciesColorsPul[unique(data_for_reproductive$Species)], fontsize = 20), nrow = 1)

# Save the plot with the legend
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.reproductive.species_hull_legend.png", width = 800, height = 450)
# Set up an empty plot area
par(mar = c(0, 0, 0, 0))
plot.new()

# Draw the legend in the empty plot area
grid.draw(grob_legend)

# Close the plotting device
dev.off()
```


#### Strain plot
```{r}
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.reproductive.strains_hull.png", width = 600, height = 450)
plot_ellipse(mod_disp, pch = NA, ellipse = F, hull = T, label = F, segments = FALSE, cex = 1, col = SpeciesColorsPul, main = "", xlab = "", ylab = "")
title(xlab = "PCoA 1", ylab = "PCoA 2", cex.lab = 1.5)
#plot(mod_disp, pch = NA, ellipse = F, hull = FALSE, label = F, segments = FALSE, cex = 2, col = "grey")
#ordihull(mod_disp, groups = data_for_reproductive$Sample.source, col = "grey", draw = "polygon")
#ordihull(mod_disp, groups = data_for_reproductive$Species, col = SpeciesColorsPul, draw = "polygon")
#ordiellipse(mod_disp, groups = data_for_reproductive$Sample.source, col = "grey", draw = "polygon")
#ordiellipse(mod_disp, groups = data_for_reproductive$Species, col = SpeciesColorsPul, draw = "polygon")
points(mod_disp2$vectors[,1], mod_disp2$vectors[,2], pch = species_to_pch[data_for_reproductive$Species], cex = 1.5, col = adjustcolor("lightgray", alpha.f = 1)) # replicates - raw datapoints
#points(mod_disp2$centroids[,1], mod_disp2$centroids[,2], pch = species_to_pch_filled, cex = 3, bg = SpeciesColorsPul, col = "black") # Species centroids
#points(mod_disp2$centroids2[,1], mod_disp2$centroids2[,2], pch = species_to_pch_filled[strain_means$Species], cex = 2, bg = SpeciesColorsPul[strain_means$Species], col = NA) # Strain centroids
points(mod_disp2$centroids2[,1], mod_disp2$centroids2[,2], pch = species_to_pch_filled[strain_means$Species], cex = 2, bg = SpeciesColorsPul[strain_means$Species], col = NA) # Strain centroids
```


#### Replicates plot
```{r}
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.reproductive.replicates_hull.png", width = 600, height = 450)
plot(mod_disp, pch = NA, ellipse = F, hull = F, label = F, segments = FALSE, cex = 1, col = "lightgray", main = "", xlab = "", ylab = "")
title(xlab = "PCoA 1", ylab = "PCoA 2", cex.lab = 1.5)
#points(mod_disp2$vectors[,1], mod_disp2$vectors[,2], pch = species_to_pch[data_for_reproductive$Species], cex = 1.5, col = NA, bg = SpeciesColorsPul[data_for_reproductive$Species]) # raw datapoints
points(mod_disp2$vectors[,1], mod_disp2$vectors[,2], pch = species_to_pch[data_for_reproductive$Species], cex = 1.5, col = SpeciesColorsPul[data_for_reproductive$Species], bg = NA) # raw datapoints
#plot(mod_disp, pch = NA, ellipse = F, hull = FALSE, label = F, segments = FALSE, cex = 2, col = "grey")
#ordihull(mod_disp, groups = data_for_reproductive$Sample.source, col = named_colors)
ordihull(mod_disp, groups = data_for_reproductive$Species, col = SpeciesColorsPul)
#ordiellipse(mod_disp, groups = data_for_reproductive$Sample.source, col = "grey", draw = "polygon")
#ordiellipse(mod_disp, groups = data_for_reproductive$Species, col = SpeciesColorsPul, draw = "polygon")
for(source in unique(data_for_reproductive$Sample.source)) {
    idx <- which(data_for_reproductive$Sample.source == source)
    
    # Assuming there's a direct relationship where each Sample.source corresponds to one Species
    # And data_for_reproductive is sorted in the same order as mod_disp$vectors
    species_for_source <- unique(data_for_reproductive$Species[idx])
    
    if(length(species_for_source) == 1) { # Ensure each source maps to a single species
        # Retrieve the color for the current species
        col <- SpeciesColorsPul[species_for_source]
        
        # Calculate the convex hull for these samples
        if(length(idx) > 2) {  # Need at least 3 points to form a hull
            hull_points <- chull(mod_disp$vectors[idx,1], mod_disp$vectors[idx,2])
            hull_points <- c(hull_points, hull_points[1])  # Ensure the hull is closed

            # Draw the filled hull
            polygon(mod_disp$vectors[idx[hull_points], 1], mod_disp$vectors[idx[hull_points], 2], col=adjustcolor(col, alpha.f = 0.5), border = NA)
        }
    } else { echo("fualt")
        # Handle cases where a Sample.source maps to multiple species, if necessary
        # This part of the code depends on how you want to address such cases
    }
}
#points(mod_disp2$centroids2[,1], mod_disp2$centroids2[,2], pch = species_to_pch_filled[strain_means$Species], cex = 2, bg = SpeciesColorsPul[strain_means$Species], col = NA) # Strain centroids
#points(mod_disp2$centroids[,1], mod_disp2$centroids[,2], pch = species_to_pch_filled, cex = 3, bg = SpeciesColorsPul, col = "black") # Species centroids
```


### Betadisp vegetative species level
```{r}
### Change dataframe to average per strain
# For vegetative_results
vegetative_results$ID <- rownames(vegetative_results)

# For data_for_vegetative (assuming Species.Code is here)
data_for_vegetative$ID <- rownames(data_for_vegetative)

# Merge to add Species.Code to vegetative_results based on row names (now in ID)
vegetative_results_with_species <- merge(vegetative_results, data_for_vegetative[, c("ID", "Sample.source", "Species")], by = "ID", all.x = TRUE)

# Remove the 'ID' column before aggregation if it's not needed for the analysis
vegetative_results_with_species$ID <- NULL
vegetative_results$ID <- NULL
data_for_vegetative$ID <- NULL

# Aggregate the means for each species
strain_means <- aggregate(. ~ Species/Sample.source, data = vegetative_results_with_species, FUN=mean)

## vegetative traits
# Between strains
mod_disp_veg <- betadisper(vegdist(vegetative_results, method = "euclidean"), data_for_vegetative$Species)
disp_test <- permutest(mod_disp_veg, permutations=999)
print(disp_test)

# Calculate centroids of strains with my custom betadisper2 function creating centroids2
mod_disp_veg2 <- betadisper2(vegdist(vegetative_results, method = "euclidean"), data_for_vegetative$Species, data_for_vegetative$Sample.source)

# To directly inspect the mean distances for each group
means <- aggregate(mod_disp_veg$distances ~ mod_disp_veg$group, FUN=mean)
print(means)
print(avg_strain_dispersion <- mean(means$`mod_disp_veg$distances`))

# Pairwise comparison of centroids (variances) using Tukey HSD
tukey_test <- TukeyHSD(mod_disp_veg)
print(tukey_test)

# Plotting
boxplot(mod_disp_veg)

xlim_values <- c(-0.585, 0.32) #c(min(pcoa_vegetative_ape$vectors[,1]), max(pcoa_vegetative_ape$vectors[,1]))
ylim_values <- c(-0.20, 0.20)
species_to_pch <- setNames(c(0, 1, 2, 6, 5, 23, 3), levels(data_for_vegetative$Species)) 
species_to_pch_filled <- setNames(c(22, 21, 24, 25, 23, 5, 3), levels(data_for_vegetative$Species)) 


sample_source_to_species <- data_for_vegetative$Species[match(unique(data_for_vegetative$Sample.source), data_for_vegetative$Sample.source)]
#SpeciesColorsPul <- setNames(SpeciesColorsPul, levels(sample_source_to_species))

# Species plot
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.vegetative.species_hull.png", width = 600, height = 450)
plot_ellipse(mod_disp_veg, pch = NA, ellipse = F, hull = T, label = F, segments = FALSE, cex = 1, col = "lightgray", main = "", xlab = "", ylab = "")
title(xlab = "PCoA 1", ylab = "PCoA 2", cex.lab = 1.5)
#plot(mod_disp_veg$centroids[,1], mod_disp_veg$centroids[,2], xlim = xlim_values, ylim = ylim_values)
points(mod_disp_veg2$vectors[,1], mod_disp_veg2$vectors[,2], pch = species_to_pch[data_for_vegetative$Species], cex = 1.5, col = "lightgray") # replicates - raw datapoints
#plot(mod_disp_veg, pch = NA, ellipse = F, hull = FALSE, label = F, segments = FALSE, cex = 2, col = "grey")
#points(mod_disp_veg2$centroids2[,1], mod_disp_veg2$centroids2[,2], pch = species_to_pch_filled[strain_means$Species], cex = 2, bg = "lightgray", col = NA) # strain centroids
#ordihull(mod_disp_veg, groups = data_for_vegetative$Species, col = "grey", draw = "polygon")
points(mod_disp_veg2$centroids[,1], mod_disp_veg2$centroids[,2], pch = species_to_pch_filled, cex = 3, bg = SpeciesColorsPul, col = NA) #species centroids

```


#### Strain plot
```{r}
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.vegetative.strains_hull.png", width = 600, height = 450)
plot_ellipse(mod_disp_veg, pch = NA, ellipse = F, hull = T, label = F, segments = FALSE, cex = 1, col = SpeciesColorsPul, main = "", xlab = "", ylab = "")
title(xlab = "PCoA 1", ylab = "PCoA 2", cex.lab = 1.5)
#plot(mod_disp_veg, pch = NA, ellipse = F, hull = FALSE, label = F, segments = FALSE, cex = 2, col = "grey")
#ordihull(mod_disp_veg, groups = data_for_vegetative$Sample.source, col = "grey", draw = "polygon")
#ordihull(mod_disp_veg, groups = data_for_vegetative$Species, col = SpeciesColorsPul, draw = "polygon")
#ordiellipse(mod_disp_veg, groups = data_for_vegetative$Sample.source, col = "grey", draw = "polygon")
#ordiellipse(mod_disp_veg, groups = data_for_vegetative$Species, col = SpeciesColorsPul, draw = "polygon")
points(mod_disp_veg2$vectors[,1], mod_disp_veg2$vectors[,2], pch = species_to_pch[data_for_vegetative$Species], cex = 1.5, col = adjustcolor("lightgray", alpha.f = 1)) # replicates - raw datapoints
#points(mod_disp_veg2$centroids[,1], mod_disp_veg2$centroids[,2], pch = species_to_pch_filled, cex = 3, bg = SpeciesColorsPul, col = "black") # Species centroids
points(mod_disp_veg2$centroids2[,1], mod_disp_veg2$centroids2[,2], pch = species_to_pch_filled[strain_means$Species], cex = 2, bg = SpeciesColorsPul[strain_means$Species], col = NA) # Strain centroids
```


#### Replicates plot
```{r}
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.vegetative.replicates_hull.png", width = 600, height = 450)
plot_ellipse(mod_disp_veg, pch = NA, ellipse = F, hull = T, label = F, segments = FALSE, cex = 1, col = SpeciesColorsPul, main = "", xlab = "", ylab = "")
title(xlab = "PCoA 1", ylab = "PCoA 2", cex.lab = 1.5)
points(mod_disp_veg2$vectors[,1], mod_disp_veg2$vectors[,2], pch = species_to_pch[data_for_vegetative$Species], cex = 1.5, col = SpeciesColorsPul[data_for_vegetative$Species], bg = NA) # raw datapoints
#plot(mod_disp_veg, pch = NA, ellipse = F, hull = FALSE, label = F, segments = FALSE, cex = 2, col = "grey")
#ordihull(mod_disp_veg, groups = data_for_vegetative$Sample.source, col = named_colors)
#ordihull(mod_disp_veg, groups = data_for_vegetative$Species, col = adjustcolor("lightgray", alpha.f = 0.5), draw = "polygon")
#ordiellipse(mod_disp_veg, groups = data_for_vegetative$Sample.source, col = "grey", draw = "polygon")
#ordiellipse(mod_disp_veg, groups = data_for_vegetative$Species, col = SpeciesColorsPul, draw = "polygon")
for(source in unique(data_for_vegetative$Sample.source)) {
    idx <- which(data_for_vegetative$Sample.source == source)
    
    # Assuming there's a direct relationship where each Sample.source corresponds to one Species
    # And data_for_vegetative is sorted in the same order as mod_disp_veg$vectors
    species_for_source <- unique(data_for_vegetative$Species[idx])
    
    if(length(species_for_source) == 1) { # Ensure each source maps to a single species
        # Retrieve the color for the current species
        col <- SpeciesColorsPul[species_for_source]
        
        # Calculate the convex hull for these samples
        if(length(idx) > 2) {  # Need at least 3 points to form a hull
            hull_points <- chull(mod_disp_veg$vectors[idx,1], mod_disp_veg$vectors[idx,2])
            hull_points <- c(hull_points, hull_points[1])  # Ensure the hull is closed

            # Draw the filled hull
            polygon(mod_disp_veg$vectors[idx[hull_points], 1], mod_disp_veg$vectors[idx[hull_points], 2], col=adjustcolor(col, alpha.f = 0.2), border = NA)
        }
    } else { echo("fualt")
        # Handle cases where a Sample.source maps to multiple species, if necessary
        # This part of the code depends on how you want to address such cases
    }
}
#points(mod_disp_veg2$centroids2[,1], mod_disp_veg2$centroids2[,2], pch = species_to_pch_filled[strain_means$Species], cex = 2, bg = SpeciesColorsPul[strain_means$Species], col = NA) # Strain centroids
#ordiellipse(mod_disp_veg, groups = data_for_vegetative$Species, draw = "polygon", col = "black", alpha = 0.2, kind = "sd", border = NA)
#points(mod_disp_veg2$centroids[,1], mod_disp_veg2$centroids[,2], pch = species_to_pch_filled, cex = 3, bg = SpeciesColorsPul, col = "black") # Species centroids
```


## Adonis2 to test for differences between species centroids
```{r}
adonis2(vegetative_daisy ~ Species, data = data_for_vegetative, method = "euclidean")

adonis2(reproductive_daisy ~ Species, data = data_for_reproductive, method = "euclidean")
```


## Hypothesis  testing using multiple regression on distance matrices approach
```{r}
### Reproductive traits
# Create one matrix per category

# Species distance matrix
Z <- model.matrix(
	~ 0 + Species,
	data = data_for_reproductive
)
species_dist_reproductive <- 1 - Z %*% t(Z)
species_dist_reproductive <- lower(species_dist_reproductive) # extract  the diagonal

# Strain dist_reproductiveance matrix
Z <- model.matrix(
	~ 0 + Sample.source,
	data = data_for_reproductive
)
strain_dist_reproductive <- 1 - Z %*% t(Z)
strain_dist_reproductive <- lower(strain_dist_reproductive)

# Shelf level dist_reproductiveance matrix
Z <- model.matrix(
	~ 0 + shelfxlevel,
	data = data_for_reproductive
)
shelf_dist_reproductive <- 1 - Z %*% t(Z)
shelf_dist_reproductive <- lower(shelf_dist_reproductive)

# Shelf level dist_reproductiveance matrix
Z <- model.matrix(
	~ 0 + shelf,
	data = data_for_reproductive
)
shelfID_dist_reproductive <- 1 - Z %*% t(Z)
shelfID_dist_reproductive <- lower(shelfID_dist_reproductive)

# Shelf level dist_reproductiveance matrix
Z <- model.matrix(
	~ 0 + level,
	data = data_for_reproductive
)
level_dist_reproductive <- 1 - Z %*% t(Z)
level_dist_reproductive <- lower(level_dist_reproductive)

# Water content distance matrix
water_content <- data_for_reproductive$Flush1Fruitbody_water_content
water_dist_reproductive <- dist(water_content, method = "euclidean")

# Run MRM model
#MRM_reproductive_logistic <- MRM(reproductive_daisy ~ species_dist_reproductive + strain_dist_reproductive + shelf_dist_reproductive, nperm = 1000,method = "logistic")
#MRM_reproductive <- MRM(reproductive_daisy ~ species_dist_reproductive + strain_dist_reproductive + shelf_dist_reproductive) # linear regression
MRM_reproductive_simple <- MRM(reproductive_daisy ~ species_dist_reproductive + strain_dist_reproductive + shelfID_dist_reproductive + level_dist_reproductive) # linear regression

MRM_reproductive_watercont <- MRM(reproductive_daisy ~ species_dist_reproductive + strain_dist_reproductive + shelfID_dist_reproductive + level_dist_reproductive + water_dist_reproductive) # linear regression





MRM_reproductive_complex <- MRM(reproductive_daisy ~ species_dist_reproductive + strain_dist_reproductive + shelfID_dist_reproductive + level_dist_reproductive + shelf_dist_reproductive)
MRM_reproductive_speciesonly <- MRM(reproductive_daisy ~ species_dist_reproductive)
print(MRM_reproductive_speciesonly)
MRM_reproductive_strainonly <- MRM(reproductive_daisy ~ strain_dist_reproductive)
print(MRM_reproductive_strainonly)
print(MRM_reproductive_simple)
print(MRM_reproductive_complex)
#print(MRM_reproductive_logistic)

### Vegetative traits
# Create one matrix per category

# Species distance matrix
Z <- model.matrix(
	~ 0 + Species,
	data = data_for_vegetative
)
species_dist_vegetative <- 1 - Z %*% t(Z)
species_dist_vegetative <- lower(species_dist_vegetative)

# Strain dist_vegetativeance matrix
Z <- model.matrix(
	~ 0 + Sample.source,
	data = data_for_vegetative
)
strain_dist_vegetative <- 1 - Z %*% t(Z)
strain_dist_vegetative <- lower(strain_dist_vegetative)

# Shelf level dist_vegetativeance matrix
Z <- model.matrix(
	~ 0 + shelfxlevel,
	data = data_for_vegetative
)
shelf_dist_vegetative <- 1 - Z %*% t(Z)
shelf_dist_vegetative <- lower(shelf_dist_vegetative)

# Shelf level dist_vegetativeance matrix
Z <- model.matrix(
	~ 0 + shelf,
	data = data_for_vegetative
)
shelfID_dist_vegetative <- 1 - Z %*% t(Z)
shelfID_dist_vegetative <- lower(shelfID_dist_vegetative)

# Shelf level dist_vegetativeance matrix
Z <- model.matrix(
	~ 0 + level,
	data = data_for_vegetative
)
level_dist_vegetative <- 1 - Z %*% t(Z)
level_dist_vegetative <- lower(level_dist_vegetative)

# Water content distance matrix
#water_content <- data_for_vegetative$Flush1Fruitbody_water_content
#water_dist_vegetative <- dist(water_content, method = "euclidean")

# Run MRM model
#MRM_vegetative_logistic <- MRM(vegetative_daisy ~ species_dist_vegetative + strain_dist_vegetative + shelf_dist_vegetative, nperm = 1000, method = "logistic")
#print(MRM_vegetative_logistic)
MRM_vegetative_simple <- MRM(vegetative_daisy ~ species_dist_vegetative + strain_dist_vegetative + shelfID_dist_vegetative + level_dist_vegetative) # linear regression

#MRM_vegetative_watercont <- MRM(vegetative_daisy ~ species_dist_vegetative + strain_dist_vegetative + shelfID_dist_vegetative + level_dist_vegetative + water_dist_vegetative) # linear regression


MRM_vegetative_complex <- MRM(vegetative_daisy ~ species_dist_vegetative + strain_dist_vegetative + shelfID_dist_vegetative + level_dist_vegetative + shelf_dist_vegetative)
#MRM_vegetative <- MRM(vegetative_daisy ~ species_dist_vegetative + strain_dist_vegetative + shelf_dist_vegetative) # linear regression
MRM_vegetative_speciesonly <- MRM(vegetative_daisy ~ species_dist_vegetative)
print(MRM_vegetative_speciesonly)
MRM_vegetative_strainonly <- MRM(vegetative_daisy ~ strain_dist_vegetative)
print(MRM_vegetative_strainonly)
#print(MRM_vegetative)
print(MRM_vegetative_simple)
print(MRM_vegetative_complex)


```

## MRM results table
```{r}
# Convert the matrix to a data frame
veg_df <- as.data.frame(MRM_vegetative_simple$coef)
# Set the first column as the row names from the matrix
veg_df$Description <- rownames(veg_df)
# Reorder the columns to have the description first
veg_df <- veg_df[, c("Description", "vegetative_daisy", "pval")]

MRM_effect_levels <- c("Intercept", "Level", "Shelf", "Species", "Strain")

# Renaming entries
veg_df$Description <- factor(veg_df$Description, labels = MRM_effect_levels)
# Rename the columns
names(veg_df)[names(veg_df) == "Description"] <- "Effect"
names(veg_df)[names(veg_df) == "vegetative_daisy"] <- "Regression coefficient"
names(veg_df)[names(veg_df) == "pval"] <- "p"
veg_df$type <- rep("Vegetative traits", 5)
veg_df <- veg_df[c(4, 1, 2, 3)]

#Creating table
MRM_vegetative_table <- flextable(veg_df)
MRM_vegetative_table <- set_formatter(MRM_vegetative_table, "Regression coefficient" = function(x) {
    formatC(x, format = "f", digits = 3)
  })

# Convert the matrix to a data frame
coef_df <- as.data.frame(MRM_reproductive_simple$coef)
# Set the first column as the row names from the matrix
coef_df$Description <- rownames(coef_df)
# Reorder the columns to have the description first
coef_df <- coef_df[, c("Description", "reproductive_daisy", "pval")]

# Renaming entries
coef_df$Description <- factor(coef_df$Description, labels = MRM_effect_levels)
# Rename the columns
names(coef_df)[names(coef_df) == "Description"] <- "Effect"
names(coef_df)[names(coef_df) == "reproductive_daisy"] <- "Regression coefficient"
names(coef_df)[names(coef_df) == "pval"] <- "p"
coef_df$type <- rep("Reproductive traits", 5)
coef_df <- coef_df[c(4, 1, 2, 3)]

#Creating table
MRM_reproductive_table <- flextable(coef_df)
MRM_reproductive_table <- set_formatter(MRM_reproductive_table, "Regression coefficient" = function(x) {
    formatC(x, format = "f", digits = 3)
  })

#Combined flextable
MRM_table <- flextable(rbind(veg_df, coef_df))
MRM_table <- set_formatter(MRM_table, "Regression coefficient" = function(x) {
    formatC(x, format = "f", digits = 3)
  })
MRM_table <- autofit(MRM_table)

save_as_docx(
  "MRM results" = MRM_table, 
  path = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/flextable_MRM.docx")
```



# Hypothesis 2: Wild vs cultivated

## Betadisper to visualise species and strain centroids
### Group differences plots
```{r}
## A) All traits
# Filter for cultivated/wild testing
data_for_hyp2 <- data_for_distance[!data_for_distance$Species == "Pul_intl",]
data_for_hyp2 <- data_for_hyp2[!data_for_hyp2$Species == "Dja",]
data_for_hyp2 <- data_for_hyp2[!apply(is.na(data_for_hyp2[includeTraits]), 1, any), ] # remove any rows with NAs
data_scaled_hyp2 <- decostand(data_for_hyp2[includeTraits], method = "standardize")

# PERMANOVA
mod_perm <- adonis2(vegdist(data_scaled_hyp2, method = "euclidean") ~ Cultivated.wild, data = data_for_hyp2, permutations = 999)
print(mod_perm)

# PERMDISP to test for differences in dispersion
mod_disp <- betadisper(vegdist(data_scaled_hyp2, method = "euclidean"), data_for_hyp2$Cultivated.wild)
disp_test <- permutest(mod_disp, permutations=999)
print(disp_test)
boxplot(mod_disp)

# To directly inspect the mean distances for each group
means <- aggregate(mod_disp$distances ~ mod_disp$group, FUN=mean)
print(means)

pch_cultwild <- c(24, 22, 21)
col_cultwild <- c("#6B7900", "#9C0039")
col_cultwild <- c("lightgray", "black")

# Second model to plot strain centroids
strain_means <- aggregate(. ~ Species/Sample.source, data = data_for_hyp2, FUN=mean) # Aggregate the means for each species to plot later
strain_means <- strain_means[-8]
distinct_data <- distinct(data_for_hyp2, Sample.source, Cultivated.wild)
strain_means <- left_join(strain_means, distinct_data, by = "Sample.source")
mod_disp2 <- betadisper(vegdist(data_scaled_hyp2, method = "euclidean"), data_for_hyp2$Sample.source)

# Plotting
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.hyp2.betadisp.png", width = 600, height = 450)
plot(mod_disp, ellipse = F, hull = F, label = F, pch = NA, segments = FALSE, cex = 1, main = "", xlab = "", ylab = "")
title(xlab = "PCoA 1", ylab = "PCoA 2", cex.lab = 1.5)
ordihull(mod_disp, label = F, groups = data_for_hyp2$Cultivated.wild, col = col_cultwild, draw = "polygon", border = NA, alpha = 0.4)
#points(mod_disp2$centroids[,1], mod_disp2$centroids[,2], pch = pch_cultwild[data_for_hyp2$Cultivated.wild], cex = 3, col = "black", bg = SpeciesColorsPul[data_for_hyp2$Species]) # replicates - raw datapoints
#ordihull(mod_disp, label = F, groups = data_for_hyp2$Sample.source, col = SpeciesColorsPul[data_for_hyp2$Cultivated.wild])
#ordihull(mod_disp, label = F, groups = data_for_hyp2$Species, col = SpeciesColorsPul, draw = "polygon", border = NA, alpha = 0.5)
for(source in unique(data_for_hyp2$Sample.source)) {
    idx <- which(data_for_hyp2$Sample.source == source)
    
    # Assuming there's a direct relationship where each Sample.source corresponds to one Species
    # And data_for_hyp2 is sorted in the same order as mod_disp$vectors
    species_for_source <- unique(data_for_hyp2$Species[idx])
    
    if(length(species_for_source) == 1) { # Ensure each source maps to a single species
        # Retrieve the color for the current species
        col <- SpeciesColorsPul[species_for_source]
        
        # Calculate the convex hull for these samples
        if(length(idx) > 2) {  # Need at least 3 points to form a hull
            hull_points <- chull(mod_disp$vectors[idx,1], mod_disp$vectors[idx,2])
            hull_points <- c(hull_points, hull_points[1])  # Ensure the hull is closed

            # Draw the filled hull
            polygon(mod_disp$vectors[idx[hull_points], 1], mod_disp$vectors[idx[hull_points], 2], col=adjustcolor(col, alpha.f = 0.6), border = NA)
        }
    } else { echo("fualt")
        # Handle cases where a Sample.source maps to multiple species, if necessary
        # This part of the code depends on how you want to address such cases
    }
}
points(mod_disp$vectors[,1], mod_disp$vectors[,2], pch = pch_cultwild[data_for_hyp2$Cultivated.wild], cex = 1.3, col = "black", bg = SpeciesColorsPul[data_for_hyp2$Species]) # replicates - raw datapoints
dev.off()


# Plotting strain centroids
png("C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.hyp2.betadisp.strains.png", width = 600, height = 450)
plot(mod_disp, ellipse = F, hull = F, label = F, pch = NA, segments = FALSE, cex = 1, main = "", xlab = "", ylab = "")
title(xlab = "PCoA 1", ylab = "PCoA 2", cex.lab = 1.5)
ordihull(mod_disp, label = F, groups = data_for_hyp2$Cultivated.wild, col = col_cultwild, draw = "polygon", border = NA, alpha = 0.4)
points(mod_disp$vectors[,1], mod_disp$vectors[,2], pch = pch_cultwild[data_for_hyp2$Cultivated.wild], cex = 1.3, col = "black") # replicates - raw datapoints
points(mod_disp2$centroids[,1], mod_disp2$centroids[,2], pch = pch_cultwild[strain_means$Cultivated.wild], cex = 3, bg = SpeciesColorsPul[strain_means$Species]) # strain centroids
#ordispider(mod_disp2, label = F, groups = data_for_hyp2$Sample.source, col = SpeciesColorsPul[strain_means$Species], lwd = 3)

# Plotting type legend
legend("bottomleft", legend = c("Cultivated", "Wild"), pch = c(22, 21), col = NA, title = "Type", cex = 1.5, bty = "n", bg = NA, pt.bg = c("lightgray", "darkgray"), pt.cex = 3)

# Plotting species legend
legend("bottomright", legend = c("P. parsonsiae", "P. pulmonarius 'NZ'"), col = NA, fill = SpeciesColorsPul[strain_means$Species], title = "Species", cex = 1.5, bty = "n", bg = NA)
```


## glmmTMB per trait - for between strain variation (nested within species) is lower in cultivated strains than in wild strains
```{r}
## All traits
# PCA on all traits
pca_all_traits <- prcomp(data_scaled_hyp2, scale=T)
print(biplot(pca_all_traits))
summary(pca_all_traits)

# nested mixed effects model, using group as a fixed effect
alltraits_hyp2_pca1_model <- glmmTMB(pca_all_traits$x[,1] ~ (1 | Species/Sample.source)+ (0 + Cultivated.wild | Species/Sample.source) + (1 | shelf * level), data = data_for_hyp2, family = gaussian())
summary(alltraits_hyp2_pca1_model)

alltraits_hyp2_pca1_reduced <- glmmTMB(pca_all_traits$x[,1] ~ (1 | Species/Sample.source) + (1 | shelf * level), data = data_for_hyp2, family = gaussian())
summary(alltraits_hyp2_pca1_reduced)
print(anova(alltraits_hyp2_pca1_model, alltraits_hyp2_pca1_reduced, test = "LRT"))

alltraits_hyp2_pca1_model_lme4 <- lmer(pca_all_traits$x[,1] ~ (1 + Cultivated.wild | Species/Sample.source) + (1 | shelf:level), data = data_for_hyp2)
summary(alltraits_hyp2_pca1_model_lme4)

alltraits_hyp2_pca1_reduced_lme4 <- lmer(pca_all_traits$x[,1] ~ (1 | Species/Sample.source) + (1 | shelf:level), data = data_for_hyp2)
summary(alltraits_hyp2_pca1_reduced_lme4)

print(anova(alltraits_hyp2_pca1_model_lme4, alltraits_hyp2_pca1_reduced_lme4))

alltraits_hyp2_pca2_model <- glmmTMB(pca_all_traits$x[,2] ~ Cultivated.wild + (1 | Species/Sample.source) + (1 | shelfxlevel), data = data_for_hyp2, family = gaussian())
summary(alltraits_hyp2_pca2_model)

# Running one model per trait, per group
### Cultivated model
# Initialize an empty list to store the model results
model_results <- list()

# Loop over each trait in the 'includeTraits' vector
for (trait in includeTraits) {
  # Create the formula dynamically using the current trait
  formula <- as.formula(paste(trait, "~ (1 | Species/Sample.source) + (1 | shelf * level)"))

  # Run the glmmTMB model for the current trait
  model <- glmmTMB(formula, data = data_for_hyp2[data_for_hyp2$Cultivated.wild == "cultivated",], family = gaussian())

  # Store the summary of the model in the list
  model_results[[trait]] <- summary(model)
}
model_results_cultivated <- model_results

# To convert the list of model summaries to a dataframe 
results_dataframe <- data.frame(
  Trait = includeTraits,
  #Trait = vegetative_traits,
  #Trait = reproductive_traits,
  Intercept_Estimate = sapply(model_results, function(x) x$coefficients$cond["(Intercept)", "Estimate"]),
  Intercept_StdError = sapply(model_results, function(x) x$coefficients$cond["(Intercept)", "Std. Error"]),
  Intercept_pValue = sapply(model_results, function(x) x$coefficients$cond["(Intercept)", "Pr(>|z|)"]),
#  Cultivated_Wild_Estimate = sapply(model_results, function(x) x$coefficients$cond["Cultivated.wildwild", "Estimate"]),
#  Cultivated_Wild_StdError = sapply(model_results, function(x) x$coefficients$cond["Cultivated.wildwild", "Std. Error"]),
#  Cultivated_Wild_pValue = sapply(model_results, function(x) x$coefficients$cond["Cultivated.wildwild", "Pr(>|z|)"]),
  Variance_Sample_Source_Species = sapply(model_results, function(x) x$varcor$cond$`Sample.source:Species`[1, 1]),
  StdDev_Sample_Source_Species = sapply(model_results, function(x) sqrt(x$varcor$cond$`Sample.source:Species`[1, 1])),
  Variance_Species = sapply(model_results, function(x) x$varcor$cond$Species[1, 1]),
  StdDev_Species = sapply(model_results, function(x) sqrt(x$varcor$cond$Species[1, 1])),
  Variance_Shelf = sapply(model_results, function(x) x$varcor$cond$shelf[1, 1]),
  StdDev_Shelf = sapply(model_results, function(x) sqrt(x$varcor$cond$shelf[1, 1])),
  Variance_Level = sapply(model_results, function(x) x$varcor$cond$level[1, 1]),
  StdDev_Level = sapply(model_results, function(x) sqrt(x$varcor$cond$level[1, 1])),
  Variance_Shelf_Level = sapply(model_results, function(x) x$varcor$cond$`shelf:level`[1, 1]),
  StdDev_Shelf_Level = sapply(model_results, function(x) sqrt(x$varcor$cond$`shelf:level`[1, 1])),
  Residual_Variance = sapply(model_results, function(x) x$sigma^2)
)

results_dataframe_cultivated <- results_dataframe

### Wild model
# Initialize an empty list to store the model results
model_results <- list()

# Loop over each trait in the 'includeTraits' vector
for (trait in includeTraits) {
  # Create the formula dynamically using the current trait
  formula <- as.formula(paste(trait, "~ (1 | Species/Sample.source) + (1 | shelf * level)"))

  # Run the glmmTMB model for the current trait
  model <- glmmTMB(formula, data = data_for_hyp2[data_for_hyp2$Cultivated.wild == "wild",], family = gaussian())

  # Store the summary of the model in the list
  model_results[[trait]] <- summary(model)
}
model_results_wild <- model_results

# To convert the list of model summaries to a dataframe 
results_dataframe <- data.frame(
  Trait = includeTraits,
  #Trait = vegetative_traits,
  #Trait = reproductive_traits,
  Intercept_Estimate = sapply(model_results, function(x) x$coefficients$cond["(Intercept)", "Estimate"]),
  Intercept_StdError = sapply(model_results, function(x) x$coefficients$cond["(Intercept)", "Std. Error"]),
  Intercept_pValue = sapply(model_results, function(x) x$coefficients$cond["(Intercept)", "Pr(>|z|)"]),
#  Cultivated_Wild_Estimate = sapply(model_results, function(x) x$coefficients$cond["Cultivated.wildwild", "Estimate"]),
#  Cultivated_Wild_StdError = sapply(model_results, function(x) x$coefficients$cond["Cultivated.wildwild", "Std. Error"]),
#  Cultivated_Wild_pValue = sapply(model_results, function(x) x$coefficients$cond["Cultivated.wildwild", "Pr(>|z|)"]),
  Variance_Sample_Source_Species = sapply(model_results, function(x) x$varcor$cond$`Sample.source:Species`[1, 1]),
  StdDev_Sample_Source_Species = sapply(model_results, function(x) sqrt(x$varcor$cond$`Sample.source:Species`[1, 1])),
  Variance_Species = sapply(model_results, function(x) x$varcor$cond$Species[1, 1]),
  StdDev_Species = sapply(model_results, function(x) sqrt(x$varcor$cond$Species[1, 1])),
  Variance_Shelf = sapply(model_results, function(x) x$varcor$cond$shelf[1, 1]),
  StdDev_Shelf = sapply(model_results, function(x) sqrt(x$varcor$cond$shelf[1, 1])),
  Variance_Level = sapply(model_results, function(x) x$varcor$cond$level[1, 1]),
  StdDev_Level = sapply(model_results, function(x) sqrt(x$varcor$cond$level[1, 1])),
  Variance_Shelf_Level = sapply(model_results, function(x) x$varcor$cond$`shelf:level`[1, 1]),
  StdDev_Shelf_Level = sapply(model_results, function(x) sqrt(x$varcor$cond$`shelf:level`[1, 1])),
  Residual_Variance = sapply(model_results, function(x) x$sigma^2)
)

results_dataframe_wild <- results_dataframe

```

#### Checking for outliers
```{r}
plot(data_for_hyp2$Sample.source, data_for_hyp2$Flush1Time_to_Mature)
plot(data_for_hyp2$Sample.source, data_for_hyp2$Flush1Fruitbody_water_content)

plot(data_for_hyp2$Cultivated.wild, data_for_hyp2$Flush1Fruitbody_water_content)

ggplot(data_for_hyp2, aes(x = Cultivated.wild, y = Flush1Fruitbody_water_content)) +
  geom_violin(aes(fill = Species), alpha = 0.5, position = position_dodge(width = 0.75)) +
  geom_beeswarm(aes(fill = Species), alpha = 0.5, position = position_dodge(width = 0.75)) +
  labs(
    x = "Cultivated/Wild"
  ) +
  theme_minimal()

ggplot(data_for_hyp2, aes(x = Cultivated.wild, y = Flush1Time_to_Mature)) +
  geom_violin(aes(fill = Species), alpha = 0.5, position = position_dodge(width = 0.75)) +
  geom_beeswarm(aes(fill = Species), alpha = 0.5, position = position_dodge(width = 0.75)) +
  labs(
    x = "Cultivated/Wild"
  ) +
  theme_minimal()

# Calculate standard deviation for each unique Sample.source
sd_result <- data_for_hyp2 %>%
  #filter(Sample.source != "DH842") %>%
  group_by(Species, Cultivated.wild) %>%
  summarise(
    sd_Flush1Fruitbody_water_content = sd(Flush1Fruitbody_water_content, na.rm = TRUE),
    sd_Flush1Time_to_Mature = sd(Flush1Time_to_Mature, na.rm = TRUE)
  )

# Print the result
print(sd_result)

plot(sd_result$Cultivated.wild, sd_result$sd_Flush1Fruitbody_water_content)
plot(sd_result$Cultivated.wild, sd_result$sd_Flush1Time_to_Mature)

ggplot(sd_result, aes(x = Cultivated.wild, y = sd_Flush1Fruitbody_water_content)) +
  geom_violin(aes(fill = Species), alpha = 0.5, position = position_dodge(width = 0.75)) +
  geom_beeswarm() +
  labs(
    x = "Cultivated/Wild"
  ) +
  theme_minimal()


```


#### Results with logical test
```{r}
# Function to extract results and perform F-test with absolute variances
create_results_dataframe <- function(model_results_wild, model_results_cultivated) {
  # Initialize an empty data frame
  results_dataframe <- data.frame(
    trait = character(),
    Between_strain_variance_wild = numeric(),
    Between_strain_variance_cultivated = numeric(),
    hyp2 = logical(),  # New column for hypothesis testing results
    stringsAsFactors = FALSE
  )
  
  # Iterate over each trait in the model results
  trait_names <- names(model_results_wild)
  for (trait in trait_names) {
    # Extract variance for wild group
    var_wild <- attr(model_results_wild[[trait]]$varcor$cond$`Sample.source:Species`, "stddev")^2
    
    # Extract variance for cultivated group
    var_cultivated <- attr(model_results_cultivated[[trait]]$varcor$cond$`Sample.source:Species`, "stddev")^2
    
    # Determine if wild variance is greater than cultivated variance
    hyp2_result <- var_wild > var_cultivated
    
    # Append results to the data frame
    results_dataframe <- rbind(results_dataframe, data.frame(
      trait = trait,
      Between_strain_variance_wild = var_wild,
      Between_strain_variance_cultivated = var_cultivated,
      hyp2 = hyp2_result  # Add the logical test result to the data frame
    ))
  }
  
  return(results_dataframe)
}


# Assuming model_results_wild and model_results_cultivated are already loaded


final_results_dataframe_absolute <- create_results_dataframe(model_results_wild, model_results_cultivated)

ft_hyp2logical <- flextable(final_results_dataframe_absolute)

ft_hyp2logical <- set_formatter(ft_hyp2logical, Between_strain_variance_wild = function(x) {
    formatC(x, format = "e", digits = 2)
  })
ft_hyp2logical <- set_formatter(ft_hyp2logical, Between_strain_variance_cultivated = function(x) {
    formatC(x, format = "e", digits = 2)
  })
ft_hyp2logical <- autofit(ft_hyp2logical)

save_as_docx(
  "my table 1" = ft_hyp2logical,  
  path = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/flextable.docx")
```

### Confidence intervals of wild and cultivated strains
```{r}
# Sample sizes
n_wild <- 9
n_cultivated <- 7

# Confidence level
alpha <- 0.05

# Chi-squared critical values
A_wild <- qchisq(alpha / 2, df = n_wild - 1)
B_wild <- qchisq(1 - alpha / 2, df = n_wild - 1)
A_cultivated <- qchisq(alpha / 2, df = n_cultivated - 1)
B_cultivated <- qchisq(1 - alpha / 2, df = n_cultivated - 1)

# Function to calculate confidence intervals
calculate_ci <- function(variance, n, A, B) {
  lower_bound <- ((n - 1) * variance) / B
  upper_bound <- ((n - 1) * variance) / A
  return(c(lower_bound, upper_bound))
}

# Calculate confidence intervals for each trait
final_results_dataframe_CI <- final_results_dataframe_absolute %>%
  rowwise() %>%
  mutate(
    CI_wild_lower = calculate_ci(Between_strain_variance_wild, n_wild, A_wild, B_wild)[1],
    CI_wild_upper = calculate_ci(Between_strain_variance_wild, n_wild, A_wild, B_wild)[2],
    CI_cultivated_lower = calculate_ci(Between_strain_variance_cultivated, n_cultivated, A_cultivated, B_cultivated)[1],
    CI_cultivated_upper = calculate_ci(Between_strain_variance_cultivated, n_cultivated, A_cultivated, B_cultivated)[2]
  )

# Print the dataframe with confidence intervals
print(final_results_dataframe_CI)

# Function to check if two intervals overlap
check_overlap <- function(lower1, upper1, lower2, upper2) {
  return(!(upper1 < lower2 | upper2 < lower1))
}

# Add the Overlap column to the dataframe
final_results_dataframe_CI <- final_results_dataframe_CI %>%
  rowwise() %>%
  mutate(Overlap = check_overlap(CI_wild_lower, CI_wild_upper, CI_cultivated_lower, CI_cultivated_upper))

# Print the updated dataframe
print(final_results_dataframe_CI)


exp_sup <- function(x, digits = 2) {
  base <- format(x / 10^floor(log10(abs(x))), nsmall = digits, digits = digits, scientific = FALSE)
  exponent <- floor(log10(abs(x)))
  sprintf("%s $\\times$ 10^%d^", base, exponent)
}

# Define the function to add Markdown line break before brackets
add_linebreak <- function(var, lower, upper, digits = 2) {
  paste0(exp_sup(var, digits), "  \n(", exp_sup(lower, digits), " - ", exp_sup(upper, digits), ")")
}

# Create the publication-ready dataframe with line breaks
hyp2_results_print <- final_results_dataframe_CI %>%
  rowwise() %>%
  mutate(
    wild = add_linebreak(Between_strain_variance_wild, CI_wild_lower, CI_wild_upper, digits = 2),
    cultivated = add_linebreak(Between_strain_variance_cultivated, CI_cultivated_lower, CI_cultivated_upper, digits = 2)
  ) %>%
  select(trait, wild, cultivated, Overlap, hyp2)

# Create a named vector for mapping
trait_label_map <- setNames(traitLabelsNoUnits, includeTraits)

# Replace trait labels in the hyp2_results_print dataframe
hyp2_results_print <- hyp2_results_print %>%
  mutate(trait = trait_label_map[trait])

# Display the new dataframe
print(hyp2_results_print)

# Create the flextable
ft <- flextable(hyp2_results_print) %>%
  colformat_md() %>%
  autofit()

# Print the flextable
ft

save_as_docx(
  "Hyp2 results" = ft, 
  path = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/flextable_hyp2results.docx")
```

#### Plotting glmmTMB model results
```{r}
# Ensure the traits in final_results_dataframe_CI are properly mapped
final_results_dataframe_absolute <- final_results_dataframe_absolute %>%
  mutate(trait = trait_label_map[trait])
final_results_dataframe_CI <- final_results_dataframe_CI %>%
  mutate(trait = trait_label_map[trait])

# Reshape the dataframe for plotting
plot_data <- final_results_dataframe_CI %>%
  pivot_longer(cols = c("Between_strain_variance_wild", "Between_strain_variance_cultivated"), 
               names_to = "Population", 
               values_to = "Variance") %>%
  mutate(Population = ifelse(grepl("wild", Population), "Wild", "Cultivated")) %>%
  pivot_wider(names_from = Population, values_from = Variance)

# Create separate dataframes for wild and cultivated confidence intervals
ci_wild <- final_results_dataframe_CI %>%
  select(trait, CI_wild_lower, CI_wild_upper) %>%
  rename(lower = CI_wild_lower, upper = CI_wild_upper) %>%
  mutate(Population = "Wild")

ci_cultivated <- final_results_dataframe_CI %>%
  select(trait, CI_cultivated_lower, CI_cultivated_upper) %>%
  rename(lower = CI_cultivated_lower, upper = CI_cultivated_upper) %>%
  mutate(Population = "Cultivated")

# Combine confidence interval dataframes
ci_data <- rbind(ci_wild, ci_cultivated)

# Merge with plot_data
plot_data <- plot_data %>%
  pivot_longer(cols = c("Wild", "Cultivated"), names_to = "Population", values_to = "Variance") %>%
  left_join(ci_data, by = c("trait", "Population"))

# Create alternating background colors for each trait
plot_data <- plot_data %>%
  mutate(trait = factor(trait, levels = unique(final_results_dataframe_absolute$trait)),
         trait_index = as.numeric(trait),
         background_color = ifelse(trait_index %% 2 == 0, "#E6E6E6", "darkgray"))

# Calculate the y-range for the plot
y_range <- range(plot_data$Variance, plot_data$lower, plot_data$upper, na.rm = TRUE)

# Create a dataframe for background rectangles
rect_data <- plot_data %>%
  distinct(trait, trait_index, background_color) %>%
  mutate(xmin = trait_index - 0.5, xmax = trait_index + 0.5, ymin = y_range[1], ymax = y_range[2])

# Verify rect_data
print(rect_data)

# Create the plot with alternating background colors
hyp2_ci_points <- ggplot() +
  geom_rect(data = rect_data, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = background_color), alpha = 0.3) +
  geom_point(data = plot_data, aes(x = factor(trait_index), y = Variance, color = Population), position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(data = plot_data, aes(x = factor(trait_index), ymin = lower, ymax = upper, color = Population), position = position_dodge(width = 0.5), width = 0.25) +
  scale_y_log10() +
  scale_x_discrete(labels = unique(plot_data$trait)) +
  scale_fill_identity() +
  labs(x = "Trait", y = "Absolute Variance (log scale)", color = "Population") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, margin = margin(t = 1)),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
print(hyp2_ci_points)

ggsave(filename = "C:/Users/herad/OneDrive - MWLR/Fungi, Food and Fodder/05 - Experiment 1 Trait Variation all strains/plot.hyp2_traits.png", plot = hyp2_ci_points, width = 10, height = 6, dpi = 300, bg = "white")
```
